### 2017-05-23 paper/paper.tex
Add start of paper.

\title{XNOR}
\author{Maciej Kula}
\date{\today}

\documentclass[sigconf]{acmart}
\begin{document}

\begin{abstract}
  "Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum."
\end{abstract}

\maketitle

\section{Introduction}

Industry ranking and recommendation systems need to scale to tens or hundreds of millions of items and users. Ranking latency (in on-line settings) and compute and storage costs (in pre-compute settings) place significant contraints on the design of the resulting system.

To alleviate these constraints, we propose the use of binary item and user representations in learning-to-rank matrix factorization models. Following the approach of [XNOR], we show that we can estimate binary user and item representations that achieve comparable accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory. In our MovieLens 100k experiments, our binary representations are within 1\% of the accuracy of the real-valued model, but are over 2 times faster to score and take 50\% of the memory.

\section{Related work}
\subsection{Heuristics}
\subsection{Approximate nearest neighbours}
\subsection{Quantization}

\section{Our approach}
Note that the binary dot produt approach generalizes beyond simple factorization models, and can be applied as a component of any model whose final scoring step involves a dot product between user and item representations. In particular, models using recurrent (like Session-based Recommendations with Recurrent Neural Networks) or convolutional (Images Don’t Lie: Transferring Deep Visual Semantic
Features to Large-Scale Multimodal Learning to Rank) item or user representations can easily be augmented to use binary dot products in the final ranking stages.

The amount of computation required for recommendations, as well the memory used to store item and user representations,

 imposes significant constraints on the design of the ranking system. In on-line recommendations, latency constraints force the use of heuristics

We use the approach developed by XNOR-Net to approximate latent-space dot products through binary operations. In this approach, we approximate the dot product between the user vector $\vec{u_i}$ and the item vector $\vec{i_i}$ by
\begin{equation}
\mathrm{sign}(\vec{u_i}) \cdot \mathrm{sign}(\vec{i_i}) \cdot \Vert\vec{u_i}\Vert_{l1} \cdot \Vert\vec{i_i}\Vert_{l1}
\end{equation}

\section{Implementation}
The model is implemented in PyTorch and trained using the nVidia K40 GPUs. During training, the embedding parameters are stored as floats. Similarly, the XNOR dot product is carried out using floats in $\{1, -1\}$. The minibatch size and number of training epochs are treated as model hyperparameters. All models are trained using the Adam training rate schedule.

The prediction code runs on the CPU and is implemented using Intel X86 AVX2 SIMD intrinsics. SIMD (Single Instruction Multiple Data) instructions allow the CPU to operate on multiple pieces of data in parallel, achieving significant speedups over the scalar version. We use explicit intrinsics rather than compiler autovectorization to ensure that neither the real-valued nor the binary prediction code is unfairly disadvantaged by the quality of compiler autovectorization.

The real-valued prediction code is implemented using 8-float wide fused multiply-add instructions (\texttt{\_mm256\_fmadd\_ps}).

In the binary version, the 1-bit weights are packed into 32-bit integer buffers. The XNOR dot product is implemented using 8-integer wide XOR operations (256 binary weights are processed at a time), followed by a popcount instruction to count the number of on bits in the result. We use the X algorithm, as provided by the Y library, for implementing the bit counting.

Both versions use 32-bit aligned input data to utilise aligned SIMD register load instructions.

The code is compiled using GCC X for the AVX2-enabled Skylake architecture.

%\bibliographystyle{abbrv}
%\bibliography{main}

\end{document}

### 2017-05-25 paper/paper.tex
Some updates.

@@
-\title{XNOR}
+\title{Binary Latent Representations for Efficient Ranking}
 \author{Maciej Kula}
 \date{\today}
+
+% --- Symbols
+\newcommand\symUserSet{U}
+\newcommand\symItemSet{I}
+\newcommand\symUserFeaturesSet{F^U}
+\newcommand\symItemFeaturesSet{F^I}
+\newcommand\symItemEmbeddingMatrix{E^I}
+\newcommand\symUserEmbeddingMatrix{E^U}
+
+\newcommand\symUserInteractionSet{S}
 
@@
 To alleviate these constraints, we propose the use of binary item and user representations in learning-to-rank matrix factorization models.
-Following the approach of [XNOR], we show that we can estimate binary user and item representations that achieve comparable accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory.
+Following the approach of \cite{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory.
 In our MovieLens 100k experiments, our binary representations are within 1\% of the accuracy of the real-valued model, but are over 2 times faster to score and take 50\% of the memory.
@@
 \subsection{Heuristics}
-\subsection{Approximate nearest neighbours}
+\subsection{Approximate inner product search}
+\cite{shrivastava2014asymmetric} propose a solution fo efficient ranking of large result sets through approximate maximum inner product search (MIPS).
+They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmtric transformations of the query and candidate vectors, and show their applicability to recommender systems.
+\cite{Shen_2015_ICCV} expand on this approach by learning asymmetric binary hash functions.
+The hash functions are trained to minimize the L2 norm between the inner product matrix of the original input vectors and their trained binary representations.
 \subsection{Quantization}
@@
 \section{Our approach}
+\cite{rastegari2016xnor} introduce XNOR-Networks, an approach that combines 1-bit quantization of fully-connected and convolutional layers with representing dot products through scaled XNOR and bitcounting operations.
+
+We adapt their approach to the recommendation task by learning binary versions of user and item vectors within a learning-to-rank factorization model.
+
+Formally, let $\symUserSet$ be the set of users, and $\symItemSet$ be the set of items.
+Each user interacts with a number of items $\symUserInteractionSet_i^+$; the set of all remaining items is denoted by $\symUserInteractionSet^-$.
+
+In a traditional real-valued model, the prediction $r_{ui}$ for any user-item interaction pair $(u, i) \in \symUserSet \times \symItemSet$ would be given by
+\begin{equation}
+r_{ui} = \sigma(\vec{u_i}) \cdot \vec{i_i} + b_i + b_u)
+\end{equation}
+where $\sigma$ denotes the sigmoid function, $\vec{u_i}$ the user and $\vec{i_i}$ the item latent vector.
+
+In the binary representation model the above is approximated by
+\begin{equation}
+\mathrm{sign}(\vec{u_i}) \cdot \mathrm{sign}(\vec{i_i}) \cdot \Vert\vec{u_i}\Vert_{l1} \cdot \Vert\vec{i_i}\Vert_{l1}
+\end{equation}
+
+% explain the derivation
+
+We train the model using the Bayesian Personalised Ranking (BPR, \cite{rendle2009bpr}) loss,
+
 Note that the binary dot produt approach generalizes beyond simple factorization models, and can be applied as a component of any model whose final scoring step involves a dot product between user and item representations.

### 2017-06-04 paper/paper.tex
Add bibliography entries.

@@
 They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmtric transformations of the query and candidate vectors, and show their applicability to recommender systems.
+
+Our approach is complementary to approximate inner product search.
+Approximate MIPS proceeds in two stages: (1) identification of candidate result vectors through localicty-sensitive hashing, followed by (2) explicit scoring and sorting of the returned candidates.
+Replacing real-valued dot products by XNOR operations in the second stage will yield incremental performance gains over a real-valued MIPS approach.
+
 \cite{Shen_2015_ICCV} expand on this approach by learning asymmetric binary hash functions.
@@
 Note that the binary dot produt approach generalizes beyond simple factorization models, and can be applied as a component of any model whose final scoring step involves a dot product between user and item representations.
-In particular, models using recurrent (like Session-based Recommendations with Recurrent Neural Networks) or convolutional (Images Don’t Lie: Transferring Deep Visual Semantic
-Features to Large-Scale Multimodal Learning to Rank) item or user representations can easily be augmented to use binary dot products in the final ranking stages.
+In particular, models using recurrent \citep{hidasi2015session} or convolutional \citep{lynch2015images} item or user representations can easily be augmented to use binary dot products in the final ranking stages.
 
@@
 
-\section{Implementation}
+\section{Experiments}
+To assess the accuracy-speed trade-offs enabled by the XNOR approach, we conduct an experiment on the Movielens 100K dataset \citep{harper2016movielens}.
+The dataset contains 100 thousand ratings from 943 users on 1682 movies.
+Despite its small size, the results on this dataset should be indicative of results that can be obtained on larger datasets, and its small size enables us to easily test hyperparameter configurations and estimate models with high latent dimensions.
+
+\subsection{Experimental setup}
+We randomly divide the dataset into a training, test, and validation set.
+We build real-valued and XOR models for between 32 and 4096 latent dimensions.
+For every latent dimensionality, we conduct a random search over the hyperparameter space, and pick optimal learning rates and minibatch sizes based on the test set.
+The final results are obtained from ranking interactions from the validation set.
+
+Benchmark results are measured on an Intel Xeon E5-2686v4 CPU.
+
+\subsection{Implementation}
 The model is implemented in PyTorch and trained using the nVidia K40 GPUs.
@@
 The XNOR dot product is implemented using 8-integer wide XOR operations (256 binary weights are processed at a time), followed by a popcount instruction to count the number of on bits in the result.
-We use the X algorithm, as provided by the Y library, for implementing the bit counting.
+We use the \texttt{libpopcnt}\footnote{\texttt{https://github.com/kimwalisch/libpopcnt}} \citep{mula2016faster} library for the bit counting operations.
 
@@
 
-The code is compiled using GCC X for the AVX2-enabled Skylake architecture.
+The code is compiled using GCC 5.4.0 for the AVX2-enabled Broadwell architecture.
 
-%\bibliographystyle{abbrv}
-%\bibliography{main}
+\subsection{Results}
+
+\bibliographystyle{abbrv}
+\bibliography{bibliography}
 

### 2017-06-05 paper/paper.tex
Add experimental setup section.

@@
 Note that the binary dot produt approach generalizes beyond simple factorization models, and can be applied as a component of any model whose final scoring step involves a dot product between user and item representations.
-In particular, models using recurrent \citep{hidasi2015session} or convolutional \citep{lynch2015images} item or user representations can easily be augmented to use binary dot products in the final ranking stages.
+In particular, models using recurrent (like Session-based Recommendations with Recurrent Neural Networks) or convolutional (Images Don’t Lie: Transferring Deep Visual Semantic
+Features to Large-Scale Multimodal Learning to Rank) item or user representations can easily be augmented to use binary dot products in the final ranking stages.
 
@@
 
-\bibliographystyle{abbrv}
-\bibliography{bibliography}
+%\bibliographystyle{abbrv}
+%\bibliography{main}
 

### 2017-06-07 paper/paper.tex
Notation improvements, latex table output.

--- 2017-06-04 paper/paper.tex
@@
-\title{Binary Latent Representations for Efficient Ranking}
-\author{Maciej Kula}
-\date{\today}
+\documentclass[sigconf]{acmart}
+
+\usepackage{booktabs}
+\usepackage{commath}
+
+%\usepackage{tabularx}
 
@@
 
-\documentclass[sigconf]{acmart}
+\DeclareMathOperator{\sign}{sign}
+\renewcommand\vec{\mathbf}
+
 \begin{document}
 
+\title{Binary Latent Representations for Efficient Ranking}
+\author{Maciej Kula}
+\date{\today}
+
 \begin{abstract}
-  "Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum."
+Multifrequency media access control has been well understood in
+general wireless ad hoc networks, while in wireless sensor networks,
+researchers still focus on single frequency solutions. In wireless
+sensor networks, each device is typically equipped with a single
+radio transceiver and applications adopt much smaller packet sizes
+compared to those in general wireless ad hoc networks. Hence, the
+multifrequency MAC protocols proposed for general wireless ad hoc
+networks are not suitable for wireless sensor network applications,
+which we further demonstrate through our simulation experiments. In
+this article, we propose MMSN, which takes advantage of
+multifrequency availability while, at the same time, takes into
+consideration the restrictions of wireless sensor networks. Through
+extensive experiments, MMSN exhibits the prominent ability to utilize
+parallel transmissions among neighboring nodes. When multiple physical
+frequencies are available, it also achieves increased energy
+efficiency, demonstrating the ability to work against radio
+interference and the tolerance to a wide range of measured time
+synchronization errors.
 \end{abstract}
@@
 To alleviate these constraints, we propose the use of binary item and user representations in learning-to-rank matrix factorization models.
-Following the approach of \cite{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory.
+Following the approach of \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory.
 In our MovieLens 100k experiments, our binary representations are within 1\% of the accuracy of the real-valued model, but are over 2 times faster to score and take 50\% of the memory.
@@
 \section{Related work}
-\subsection{Heuristics}
-\subsection{Approximate inner product search}
-\cite{shrivastava2014asymmetric} propose a solution fo efficient ranking of large result sets through approximate maximum inner product search (MIPS).
+\citet{shrivastava2014asymmetric} propose a solution fo efficient ranking of large result sets through approximate maximum inner product search (MIPS).
 They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmtric transformations of the query and candidate vectors, and show their applicability to recommender systems.
@@
 
-\cite{Shen_2015_ICCV} expand on this approach by learning asymmetric binary hash functions.
-The hash functions are trained to minimize the L2 norm between the inner product matrix of the original input vectors and their trained binary representations.
-\subsection{Quantization}
+\citet{Shen_2015_ICCV} expand on this approach by learning asymmetric binary hash functions.
+The hash functions are trained to minimize the L2 norm between the inner product matrix of the original input vectors and the inner product matrix of their binary representations.
+As a post-processing technique, it is a applicable to a wide range of problems.
+However, because the binary representations are obtained as a separate step, this approach may not be able to exploit the benefits in training a binary representation approach end-to-end.
 
 \section{Our approach}
-\cite{rastegari2016xnor} introduce XNOR-Networks, an approach that combines 1-bit quantization of fully-connected and convolutional layers with representing dot products through scaled XNOR and bitcounting operations.
+\citet{rastegari2016xnor} introduce XNOR-Networks, an approach that combines 1-bit quantization of fully-connected and convolutional layers with representing dot products through scaled XNOR and bitcounting operations.
 
@@
 
-In a traditional real-valued model, the prediction $r_{ui}$ for any user-item interaction pair $(u, i) \in \symUserSet \times \symItemSet$ would be given by
+In a traditional real-valued latent factor model, the prediction $r_{ui}$ for any user-item interaction pair $(u, i) \in \symUserSet \times \symItemSet$ is given by
 \begin{equation}
-r_{ui} = \sigma(\vec{u_i}) \cdot \vec{i_i} + b_i + b_u)
+r_{ui} = \sigma(\vec{u}_i \cdot \vec{i}_i + b_i + b_u)
 \end{equation}
-where $\sigma$ denotes the sigmoid function, $\vec{u_i}$ the user and $\vec{i_i}$ the item latent vector.
+where $\sigma$ denotes the sigmoid function, $\vec{u}_i$ the user and $\vec{i}_i$ the item latent vectors, and $b_u$ and $b_i$ the user and item biases.
 
-In the binary representation model the above is approximated by
+\citet{rastegari2016xnor} show that the dot product of two real-valued vectors can be approximated in the binary domain by (1) binarizing the input vectors using the sign function (so that the results lie in $\{1, -1\}$), (2) taking their dot product, and (3) scaling the result by the average magnitude of the vectors' elements:
 \begin{equation}
-\mathrm{sign}(\vec{u_i}) \cdot \mathrm{sign}(\vec{i_i}) \cdot \Vert\vec{u_i}\Vert_{l1} \cdot \Vert\vec{i_i}\Vert_{l1}
+\vec{u}_i \cdot \vec{i}_i \approx \sign(\vec{u}_i) \cdot \sign(\vec{i}_i) \cdot \sum_k\abs{u_{ik}} \cdot \sum_k\abs{i_{ik}}
+\end{equation}
+
+Applying this to the latent-factor model, the prediction for a user-item pair is given by
+\begin{equation}
+r_{ui} = \sigma\left(\sign(\vec{u}_i) \cdot \sign(\vec{i}_i) \cdot \sum_k\abs{u_{ik}} \cdot \sum_k\abs{i_{ik}} + b_i + b_u\right)
 \end{equation}
@@
 
-We train the model using the Bayesian Personalised Ranking (BPR, \cite{rendle2009bpr}) loss,
+We train the model using the Bayesian Personalised Ranking (BPR, \citet{rendle2009bpr}) loss,
 
@@
  imposes significant constraints on the design of the ranking system. In on-line recommendations, latency constraints force the use of heuristics
-
-We use the approach developed by XNOR-Net to approximate latent-space dot products through binary operations.
-In this approach, we approximate the dot product between the user vector $\vec{u_i}$ and the item vector $\vec{i_i}$ by
-\begin{equation}
-\mathrm{sign}(\vec{u_i}) \cdot \mathrm{sign}(\vec{i_i}) \cdot \Vert\vec{u_i}\Vert_{l1} \cdot \Vert\vec{i_i}\Vert_{l1}
-\end{equation}
 
@@
 
-Benchmark results are measured on an Intel Xeon E5-2686v4 CPU.
+Benchmark results are measured on an Intel Xeon E5-2686v4 CPU by running a 100 repetitions of scoring 100,000 items and averaging the results.
 
@@
 The XNOR dot product is implemented using 8-integer wide XOR operations (256 binary weights are processed at a time), followed by a popcount instruction to count the number of on bits in the result.
-We use the \texttt{libpopcnt}\footnote{\texttt{https://github.com/kimwalisch/libpopcnt}} \citep{mula2016faster} library for the bit counting operations.
+We use the \texttt{libpopcnt} \citep{mula2016faster} library for the bit counting operations.
 
@@
 
-The code is compiled using GCC 5.4.0 for the AVX2-enabled Broadwell architecture.
+The code is compiled using GCC 4.8.4 for the AVX2-enabled Broadwell architecture.
 
 \subsection{Results}
+\begin{table*}
+\centering
+\begin{tabular}{rrrrrrrr}
+\toprule
+Dimension &   MRR & Binary MRR & MRR ratio &   PPMS & Binary PPMS & PPMS ratio & Memory use ratio \\
+\midrule
+       32 & 0.096 &      0.088 &     0.919 & 89,299 &     245,644 &      2.751 &            0.091 \\
+       64 & 0.098 &      0.096 &     0.973 & 47,565 &     209,486 &      4.404 &            0.062 \\
+      128 & 0.108 &      0.091 &     0.843 & 20,502 &     156,404 &      7.629 &            0.047 \\
+      256 & 0.109 &      0.097 &     0.884 & 10,631 &     166,500 &     15.661 &            0.039 \\
+      512 & 0.112 &      0.105 &     0.938 &   5391 &     110,418 &     20.479 &            0.035 \\
+     1024 & 0.114 &      0.102 &     0.897 &   2689 &      69,349 &     25.786 &            0.033 \\
+     2048 & 0.111 &      0.109 &     0.984 &   1352 &      34,826 &     25.745 &            0.032 \\
+     4096 & 0.111 &      0.107 &     0.961 &    666 &      15,907 &     23.856 &            0.032 \\
+\bottomrule
+\end{tabular}
+\end{table*}
 
-\bibliographystyle{abbrv}
 \bibliography{bibliography}
+\bibliographystyle{ACM-Reference-Format}
 

### 2017-06-08 paper/paper.tex
Fix notation. Add sections to approach.

@@
 We adapt their approach to the recommendation task by learning binary versions of user and item vectors within a learning-to-rank factorization model.
+This allows us to use binary instead of floating point operations for computing recommendation rankings, which has significant speed and memory use advantages.
+In a real-valued $n$-dimensional factorization model, item and user representations take $4n$ bytes of memory, and computing a single prediction takes $2n$ floating point operations.
+In an equivalent binary model, $\frac{3}{32}n$ binary operations are required for each prediction (XOR, negation, and bitcounting), and the representations take $\frac{1}{8}n$ bytes of memory.
+Naive calculations suggest that, for the same latent dimensionality, a binary model would be over 20 times faster to score, and take less then 5\% of memory to store its representations.
 
-Formally, let $\symUserSet$ be the set of users, and $\symItemSet$ be the set of items.
+Naturally, using binary instead of real-valued representations entails a loss of representational power, and the resulting models can be expected to be less accurate.
+Nevertheless, binary models may still be very useful in a production system, for two reasons.
+Firstly, they may offer the system designer a more favourable trade-off between speed and accuracy than simply changing the number of latent dimensions in a real-valued model.
+If latency is the binding constraint, switching to a binary model may offer better or comparable speed with better ranking performance.
+If accuracy is the priority, a binary model may offer equivalent ranking performance while being faster to score.
+This suspicion is borne out in out experiments.
+Secondly, many production systems resort to heavy use of heuristics when selecting candidates to be scored by their ranking systems (as scoring all candidates would be infeasible).
+To the extent that these heuristics are suboptimal, replacing them with a binary representation model would improve the overall accuracy of the system.
+
+To quantify the tradeoffs involved, we construct and fit both a real-valued and a binary version of simple implicit learning-to-rank factorization model.
+We introduce the notation and the models below.
+
+Let $\symUserSet$ be the set of users, and $\symItemSet$ be the set of items.
 Each user interacts with a number of items $\symUserInteractionSet_i^+$; the set of all remaining items is denoted by $\symUserInteractionSet^-$.
@@
 \begin{equation}
-r_{ui} = \sigma(\vec{u}_i \cdot \vec{i}_i + b_i + b_u)
+r_{ui} = \vec{u}_i \cdot \vec{i}_i + b_i + b_u
 \end{equation}
-where $\sigma$ denotes the sigmoid function, $\vec{u}_i$ the user and $\vec{i}_i$ the item latent vectors, and $b_u$ and $b_i$ the user and item biases.
+where $\vec{u}_i$ denotes the user and $\vec{i}_i$ the item $n$-dimensional latent vectors, and $b_u$ and $b_i$ the user and item biases.
 
-\citet{rastegari2016xnor} show that the dot product of two real-valued vectors can be approximated in the binary domain by (1) binarizing the input vectors using the sign function (so that the results lie in $\{1, -1\}$), (2) taking their dot product, and (3) scaling the result by the average magnitude of the vectors' elements:
+\citet{rastegari2016xnor} show that the dot product of two real-valued vectors can be approximated in the binary domain by (1) binarizing the input vectors using the sign function (so that the results lie in $\{1, -1\}$), (2) taking their dot product, and (3) scaling the result by the average magnitude of the vectors' elements.
+Letting
 \begin{equation}
-\vec{u}_i \cdot \vec{i}_i \approx \sign(\vec{u}_i) \cdot \sign(\vec{i}_i) \cdot \sum_k\abs{u_{ik}} \cdot \sum_k\abs{i_{ik}}
+\alpha = \frac{1}{n}\norm{\vec{i}_{i}}_{\ell1}
 \end{equation}
-
+and
+\begin{equation}
+\beta = \frac{1}{n}\norm{\vec{u}_{i}}_{\ell1}
+\end{equation}
+represent the scaling factors, the approximation is given by
+\newcommand\binaryApproximation{ \alpha \beta\left(\sign(\vec{u}_i) \cdot \sign(\vec{i}_i)\right)}
+\begin{equation}
+\vec{u}_i \cdot \vec{i}_i \approx \binaryApproximation
+\end{equation}
 Applying this to the latent-factor model, the prediction for a user-item pair is given by
 \begin{equation}
-r_{ui} = \sigma\left(\sign(\vec{u}_i) \cdot \sign(\vec{i}_i) \cdot \sum_k\abs{u_{ik}} \cdot \sum_k\abs{i_{ik}} + b_i + b_u\right)
+r_{ui} = \sigma\left(\binaryApproximation + b_i + b_u\right)
 \end{equation}
+The model is trained using backpropagation.
+To facilitate the training of the binary model, we need a smooth approximation to the derivative of the sign function.
+Following \citet{courbariaux2016binarized}, we define
+\begin{equation}
+\frac{d\sign}{dw} = w1_{\abs{w} \leq 1}.
+\end{equation}
+we use two loss functions to train the model:
+\begin{itemize}
+\item Bayesian personalised ranking (BPR, \citet{rendle2009bpr})
+\item adaptive sampling maximum margin loss, following \citet{weston2011wsabie}.
+\end{itemize}
+For both loss functions, for any known positive user-item interaction pair $(u, i)$, we sample an implicit negative item $j \in \symUserInteractionSet^-$.
+For BPR, the loss for any such triplet is given by
+\begin{equation}
+1 - \sigma\left(r_{ui} - r_{uj}\right).
+\end{equation}
+The adaptive sampling loss is given by
+\begin{equation}
+\abs{1 - r_{ui} + r_{uj}}_{+}.
+\end{equation}
+For any $(u, i)$ pair, if the sampled negative item $j$ results in a zero loss (that is, the desired pairwise ordering is not violated), a new negative item is sampled, up to a total of $k$ attempts.
+This leads the model to perform more gradient updates in areas where its ranking performance is poorest.
 
-% explain the derivation
-
-We train the model using the Bayesian Personalised Ranking (BPR, \citet{rendle2009bpr}) loss,
-
-Note that the binary dot produt approach generalizes beyond simple factorization models, and can be applied as a component of any model whose final scoring step involves a dot product between user and item representations.
+For the purposes of this paper, we focus on a simple bilinear collaborative filtering model, and do not use any external metadata information or model the sequential nature of the data.
+However, the binary dot produt approach generalizes beyond simple factorization models, and can be applied as a component of any model whose final scoring step involves a dot product between user and item representations.
 In particular, models using recurrent \citep{hidasi2015session} or convolutional \citep{lynch2015images} item or user representations can easily be augmented to use binary dot products in the final ranking stages.
-
-The amount of computation required for recommendations, as well the memory used to store item and user representations,
-
- imposes significant constraints on the design of the ranking system. In on-line recommendations, latency constraints force the use of heuristics
+Assuming item representations can be computed in advance, even very complicated models can enjoy the efficiency advantages of using binary representations.
 
@@
 \subsection{Results}
+\label{sec:results}
 \begin{table*}

### 2017-06-08 paper/paper.tex
Start fleshing out results.

@@
 \section{Our approach}
+\label{sec:approach}
 \citet{rastegari2016xnor} introduce XNOR-Networks, an approach that combines 1-bit quantization of fully-connected and convolutional layers with representing dot products through scaled XNOR and bitcounting operations.
@@
 If accuracy is the priority, a binary model may offer equivalent ranking performance while being faster to score.
-This suspicion is borne out in out experiments.
+This suspicion is borne out in our experiments.
 Secondly, many production systems resort to heavy use of heuristics when selecting candidates to be scored by their ranking systems (as scoring all candidates would be infeasible).
@@
 \begin{equation}
-r_{ui} = \sigma\left(\binaryApproximation + b_i + b_u\right)
+r_{ui} = \binaryApproximation + b_i + b_u
 \end{equation}
+Note that the scaling factors $\alpha$ and $\beta$ need to be stored (in addition to the latent representations and biases) to compute predictions in the binary model.
+
 The model is trained using backpropagation.
@@
 \begin{equation}
-1 - \sigma\left(r_{ui} - r_{uj}\right).
+1 - \sigma\left(r_{ui} - r_{uj}\right),
 \end{equation}
+where $\sigma$ denotes the sigmoid function.
 The adaptive sampling loss is given by
@@
 
+The source code for the implementation is available on Github\footnote{\url{https://github.com/maciejkula/binge}}.
+
 \subsection{Results}
 \label{sec:results}
-\begin{table*}
+\begin{table*}[h]
+\caption{Movielens 100K results}
+\label{tb:results}
 \centering
@@
 \end{table*}
+Table \ref{tb:results} summarises the results of the Movielens 100K experiment.
+Each row presents the best MRR (mean reciprocal rank) results for both real-valued and binary models of a given dimensionality.
+MRR ratio denotes the fraction of the real-valued MRR the binary model achieves for that dimensionality; PPMS (predictions per millisecond) ratio denotes how many more predictions the binary model can compute per millisecond.
+
+Results on scoring speed and memory use broadly confirm the naive calculations from section \ref{sec:approach}, converging to over 20 times faster scoring at around 3\% of memory as latent dimensionality increases.
+As expected, for the same dimensionality a binary model achieves lower ranking accuracy. On average, the accuracy loss is around 8\%, varying between 2\% and 16\%.
+Nevertheless, the binary model's speed enables some very favourable tradeoffs
+
+When optimizing for MRR in both settings (at 1024 dimensions for the continuous and 2048 for the binary case), the XNOR representation achieves 96\% of the maximum MRR of the float representations at 13 times the query throughput and using only 6\% of the memory.
+
+When optimizing for query throughput at 32 dimensions for the continuous case, the XNOR representation (at 512 dimensions) can improve the ranking performance by achieving 110\% of the continuous MRR while being slightly faster at 1.2 times the query throughput, and using 55\% of the memory.
 

### 2017-06-08 paper/paper.tex
Add abstract.

@@
 \begin{abstract}
-Multifrequency media access control has been well understood in
-general wireless ad hoc networks, while in wireless sensor networks,
-researchers still focus on single frequency solutions. In wireless
-sensor networks, each device is typically equipped with a single
-radio transceiver and applications adopt much smaller packet sizes
-compared to those in general wireless ad hoc networks. Hence, the
-multifrequency MAC protocols proposed for general wireless ad hoc
-networks are not suitable for wireless sensor network applications,
-which we further demonstrate through our simulation experiments. In
-this article, we propose MMSN, which takes advantage of
-multifrequency availability while, at the same time, takes into
-consideration the restrictions of wireless sensor networks. Through
-extensive experiments, MMSN exhibits the prominent ability to utilize
-parallel transmissions among neighboring nodes. When multiple physical
-frequencies are available, it also achieves increased energy
-efficiency, demonstrating the ability to work against radio
-interference and the tolerance to a wide range of measured time
-synchronization errors.
+Large-scale recommender systems often face severe latency and storage constraints.
+To relax those, we train recommendation models that use binary rather than real-valued user and item representations, and show that they enable a range of attractive speed-accuracy trade-offs.
+In particular, we demonstrate 13 times faster prediction at the expense of a modest 4\% loss in ranking accuracy on the Movielens 100K dataset.
 \end{abstract}
@@
 \end{equation}
-we use two loss functions to train the model:
+
+We use two loss functions to train the model:
 \begin{itemize}
-\item Bayesian personalised ranking (BPR, \citet{rendle2009bpr})
+\item Bayesian personalised ranking (BPR, \citet{rendle2009bpr}), and
 \item adaptive sampling maximum margin loss, following \citet{weston2011wsabie}.

### 2017-06-10 paper/paper.tex
Flesh out introduction.

@@
 \usepackage{commath}
-
-%\usepackage{tabularx}
 
@@
 \author{Maciej Kula}
+\email{maciej.kula@gmail.com}
 \date{\today}
+\settopmatter{printacmref=false, printccs=false, printfolios=true}
+\setcopyright{none}
+\acmConference[]{}{}{}
 
 \begin{abstract}
-Large-scale recommender systems often face severe latency and storage constraints.
+Large-scale recommender systems often face severe latency and storage constraints at prediction time.
 To relax those, we train recommendation models that use binary rather than real-valued user and item representations, and show that they enable a range of attractive speed-accuracy trade-offs.
@@
 \section{Introduction}
-
 Industry ranking and recommendation systems need to scale to tens or hundreds of millions of items and users.
@@
 
+In an on-line setting, where recommendation latency directly impacts the user experience, system designers have to work within a strict latency budget.
+Respecting that budget often means trading off model accuracy for speed, either via smaller and less expressive models (for example, with lower-dimensionality representations) or through aggressive use of heuristics to exclude large classes of candidate items from scoring.
+
+The challenge of low-latency scoring leads many systems to rely on a pre-compute system, where recommendations are computed in advance and stored for serving at a later time.
+The primary disadvantage of such a system is its inability to react dynamically to user actions (such as new interactions that allow the recommendations to be refined) and recommendation context (location, time of day or year).
+Additionaly, while a pre-compute solution removes model prediction from the critical path, it still requires substantial investment in compute and storage resources to calculate and cache the predictions of the model.
+
 To alleviate these constraints, we propose the use of binary item and user representations in learning-to-rank matrix factorization models.
-Following the approach of \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory.
-In our MovieLens 100k experiments, our binary representations are within 1\% of the accuracy of the real-valued model, but are over 2 times faster to score and take 50\% of the memory.
+Following the approach of \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable ranking accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory.
+In MovieLens 100k experiments, our binary representations are within 4\% of the accuracy of the real-valued model, but are over 13 times faster to score and take 55\% of the memory.
 
 \section{Related work}
-\citet{shrivastava2014asymmetric} propose a solution fo efficient ranking of large result sets through approximate maximum inner product search (MIPS).
+\citet{shrivastava2014asymmetric} propose a solution for efficient ranking of large result sets through approximate maximum inner product search (MIPS).
 They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmtric transformations of the query and candidate vectors, and show their applicability to recommender systems.

### 2017-06-11 paper/paper.tex
Successfully move table.

@@
-\documentclass[sigconf]{acmart}
+\documentclass[sigchi]{acmart}
 
@@
 Respecting that budget often means trading off model accuracy for speed, either via smaller and less expressive models (for example, with lower-dimensionality representations) or through aggressive use of heuristics to exclude large classes of candidate items from scoring.
+For example, approximate nearest neighbour search techniques are used at Youtube \citep{covington2016deep} to reduce the number of candidate videos that need to be scored for each user.
 
@@
 The primary disadvantage of such a system is its inability to react dynamically to user actions (such as new interactions that allow the recommendations to be refined) and recommendation context (location, time of day or year).
+This leads to often complex architecture that mix fairly simple on-line models, computationally expensive offline models, and intermediate complexity near-line models where model updating is important \citep{amatriain2013big}.
 Additionaly, while a pre-compute solution removes model prediction from the critical path, it still requires substantial investment in compute and storage resources to calculate and cache the predictions of the model.
@@
 \section{Experiments}
-To assess the accuracy-speed trade-offs enabled by the XNOR approach, we conduct an experiment on the Movielens 100K dataset \citep{harper2016movielens}.
-The dataset contains 100 thousand ratings from 943 users on 1682 movies.
-Despite its small size, the results on this dataset should be indicative of results that can be obtained on larger datasets, and its small size enables us to easily test hyperparameter configurations and estimate models with high latent dimensions.
-
-\subsection{Experimental setup}
-We randomly divide the dataset into a training, test, and validation set.
-We build real-valued and XOR models for between 32 and 4096 latent dimensions.
-For every latent dimensionality, we conduct a random search over the hyperparameter space, and pick optimal learning rates and minibatch sizes based on the test set.
-The final results are obtained from ranking interactions from the validation set.
-
-Benchmark results are measured on an Intel Xeon E5-2686v4 CPU by running a 100 repetitions of scoring 100,000 items and averaging the results.
-
-\subsection{Implementation}
-The model is implemented in PyTorch and trained using the nVidia K40 GPUs.
-During training, the embedding parameters are stored as floats.
-Similarly, the XNOR dot product is carried out using floats in $\{1, -1\}$.
-The minibatch size and number of training epochs are treated as model hyperparameters.
-All models are trained using the Adam training rate schedule.
-
-The prediction code runs on the CPU and is implemented using Intel X86 AVX2 SIMD intrinsics.
-SIMD (Single Instruction Multiple Data) instructions allow the CPU to operate on multiple pieces of data in parallel, achieving significant speedups over the scalar version.
-We use explicit intrinsics rather than compiler autovectorization to ensure that neither the real-valued nor the binary prediction code is unfairly disadvantaged by the quality of compiler autovectorization.
-
-The real-valued prediction code is implemented using 8-float wide fused multiply-add instructions (\texttt{\_mm256\_fmadd\_ps}).
-
-In the binary version, the 1-bit weights are packed into 32-bit integer buffers.
-The XNOR dot product is implemented using 8-integer wide XOR operations (256 binary weights are processed at a time), followed by a popcount instruction to count the number of on bits in the result.
-We use the \texttt{libpopcnt} \citep{mula2016faster} library for the bit counting operations.
-
-Both versions use 32-bit aligned input data to utilise aligned SIMD register load instructions.
-
-The code is compiled using GCC 4.8.4 for the AVX2-enabled Broadwell architecture.
-
-The source code for the implementation is available on Github\footnote{\url{https://github.com/maciejkula/binge}}.
-
-\subsection{Results}
-\label{sec:results}
-\begin{table*}[h]
+\begin{table*}[htbp]
 \caption{Movielens 100K results}
@@
 \end{table*}
+To assess the accuracy-speed trade-offs enabled by the XNOR approach, we conduct an experiment on the Movielens 100K dataset \citep{harper2016movielens}.
+The dataset contains 100 thousand ratings from 943 users on 1682 movies.
+Despite its small size, the results on this dataset should be indicative of results that can be obtained on larger datasets, and its small size enables us to easily test hyperparameter configurations and estimate models with high latent dimensions.
+
+\subsection{Experimental setup}
+We randomly divide the dataset into a training, test, and validation set.
+In order to build a picture of the accuracy-speed trade-offs afforded by real-valued and binary latent models, we explicitly build models for between 32 and 4096 latent dimensions.
+For every latent dimensionality, we conduct a random search over the hyperparameter space, and pick optimal learning rates and minibatch sizes for each algorithm based on their performance on the test set.
+The final results are obtained from ranking interactions from the validation set.
+We use mean reciprocal rank (MRR) as a measure of ranking quality.
+
+Benchmark results are measured on an Intel Xeon E5-2686v4 CPU by running a 100 repetitions of scoring 100,000 items and averaging the results.
+
+\subsection{Implementation}
+The model is implemented in PyTorch and trained using the nVidia K40 GPUs.
+During training, the embedding parameters are stored as floats.
+Similarly, the XNOR dot product is carried out using floats in $\{1, -1\}$.
+The minibatch size and number of training epochs are treated as model hyperparameters.
+All models are trained using the Adam training rate schedule.
+
+The prediction code runs on the CPU and is implemented using Intel X86 AVX2 SIMD intrinsics.
+SIMD (Single Instruction Multiple Data) instructions allow the CPU to operate on multiple pieces of data in parallel, achieving significant speedups over the scalar version.
+We use explicit intrinsics rather than compiler autovectorization to ensure that neither the real-valued nor the binary prediction code is unfairly disadvantaged by the quality of compiler autovectorization.
+
+The real-valued prediction code is implemented using 8-float wide fused multiply-add instructions (\texttt{\_mm256\_fmadd\_ps}).
+
+In the binary version, the real-valued embedding parameters used in training are discarded, and the derived 1-bit weights are packed into 32-bit integer buffers.
+The XNOR dot product is implemented using 8-integer wide XOR operations (256 binary weights are processed at a time), followed by a popcount instruction to count the number of on bits in the result.
+We use the \texttt{libpopcnt} \citep{mula2016faster} library for the bit counting operations.
+
+Both versions use 32-bit aligned input data to utilise aligned SIMD register load instructions. The code is compiled using GCC 4.8.4 for the AVX2-enabled Broadwell architecture.
+
+The source code for the implementation is available on Github\footnote{\url{https://github.com/maciejkula/binge}}.
+
+\subsection{Results}
+\label{sec:results}
 Table \ref{tb:results} summarises the results of the Movielens 100K experiment.
@@
 
+\section{Conclusion}
+We show that binary user and item representations allow recommender system designers to exploit an attractive new set of speed-accuracy trade-offs.
+They are straightforward to estimate, and can easily be used in any existing model estimated using stochastic gradient descent.
+
 \bibliography{bibliography}

### 2017-06-13 paper/paper.tex
Address comments.

@@
 \usepackage{commath}
+\usepackage{threeparttable}
 
@@
 In an on-line setting, where recommendation latency directly impacts the user experience, system designers have to work within a strict latency budget.
-Respecting that budget often means trading off model accuracy for speed, either via smaller and less expressive models (for example, with lower-dimensionality representations) or through aggressive use of heuristics to exclude large classes of candidate items from scoring.
-For example, approximate nearest neighbour search techniques are used at Youtube \citep{covington2016deep} to reduce the number of candidate videos that need to be scored for each user.
+Respecting that budget often means trading off model accuracy for speed, either via smaller and less expressive models (with lower-dimensionality representations) or through aggressive use of heuristics to exclude large classes of candidate items from scoring.
+For example, approximate nearest neighbour search techniques are used at YouTube \citep{covington2016deep} to reduce the number of candidate videos that need to be scored for each user.
 
-The challenge of low-latency scoring leads many systems to rely on a pre-compute system, where recommendations are computed in advance and stored for serving at a later time.
+The challenge of low-latency scoring leads many systems to rely on a pre-compute system, where recommendations are calculated in advance and stored for serving at a later time.
 The primary disadvantage of such a system is its inability to react dynamically to user actions (such as new interactions that allow the recommendations to be refined) and recommendation context (location, time of day or year).
-This leads to often complex architecture that mix fairly simple on-line models, computationally expensive offline models, and intermediate complexity near-line models where model updating is important \citep{amatriain2013big}.
-Additionaly, while a pre-compute solution removes model prediction from the critical path, it still requires substantial investment in compute and storage resources to calculate and cache the predictions of the model.
+This leads to often complex architectures that mix fairly simple on-line models, computationally expensive offline models, and intermediate complexity near-line algorithms where model updating is important \citep{amatriain2013big}.
+Additionaly, while a pre-compute solution removes model prediction from the critical path, it still requires substantial investment in computational and storage resources to calculate and cache the predictions of the model.
 
@@
 \citet{shrivastava2014asymmetric} propose a solution for efficient ranking of large result sets through approximate maximum inner product search (MIPS).
-They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmtric transformations of the query and candidate vectors, and show their applicability to recommender systems.
+They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmetric transformations of the query and candidate vectors, and show their applicability to recommender systems.
 
@@
 \begin{equation}
-r_{ui} = \vec{u}_i \cdot \vec{i}_i + b_i + b_u
+r_{ui} = \vec{u}_u \cdot \vec{i}_i + b_u + b_i
 \end{equation}
-where $\vec{u}_i$ denotes the user and $\vec{i}_i$ the item $n$-dimensional latent vectors, and $b_u$ and $b_i$ the user and item biases.
+where $\vec{u}_u$ denotes the user and $\vec{i}_i$ the item $n$-dimensional latent vectors, and $b_u$ and $b_i$ the user and item biases.
 
@@
 \begin{equation}
-\beta = \frac{1}{n}\norm{\vec{u}_{i}}_{\ell1}
+\beta = \frac{1}{n}\norm{\vec{u}_{u}}_{\ell1}
 \end{equation}
 represent the scaling factors, the approximation is given by
-\newcommand\binaryApproximation{ \alpha \beta\left(\sign(\vec{u}_i) \cdot \sign(\vec{i}_i)\right)}
+\newcommand\binaryApproximation{ \alpha \beta\left(\sign(\vec{u}_u) \cdot \sign(\vec{i}_i)\right)}
 \begin{equation}
-\vec{u}_i \cdot \vec{i}_i \approx \binaryApproximation
+\vec{u}_u \cdot \vec{i}_i \approx \binaryApproximation
 \end{equation}
@@
 \begin{equation}
-r_{ui} = \binaryApproximation + b_i + b_u
+r_{ui} = \binaryApproximation + b_u + b_i
 \end{equation}
@@
 
-The model is trained using backpropagation.
-To facilitate the training of the binary model, we need a smooth approximation to the derivative of the sign function.
-Following \citet{courbariaux2016binarized}, we define
+The model is trained using backpropagation, and so we a smooth approximation to the derivative of the sign function to facilitate training. Following \citet{courbariaux2016binarized}, we define
 \begin{equation}
@@
 For the purposes of this paper, we focus on a simple bilinear collaborative filtering model, and do not use any external metadata information or model the sequential nature of the data.
-However, the binary dot produt approach generalizes beyond simple factorization models, and can be applied as a component of any model whose final scoring step involves a dot product between user and item representations.
+However, the binary dot product approach generalizes beyond simple factorization models, and can be applied as a component of any model whose final scoring step involves a dot product between user and item representations.
 In particular, models using recurrent \citep{hidasi2015session} or convolutional \citep{lynch2015images} item or user representations can easily be augmented to use binary dot products in the final ranking stages.
@@
 \begin{table*}[htbp]
+\begin{threeparttable}
 \caption{Movielens 100K results}
@@
 \toprule
-Dimension &   MRR & Binary MRR & MRR ratio &   PPMS & Binary PPMS & PPMS ratio & Memory use ratio \\
+Dimension &   MRR & Binary MRR & MRR ratio\tnote{1} &   PPMS\tnote{2} & Binary PPMS & PPMS ratio & Memory use ratio \\
 \midrule
@@
 \end{tabular}
+\begin{tablenotes}
+\small{
+\item[1] Ratio of binary model MRR to real-valued model MRR
+\item[2] Predictions Per Millisecond
+}
+\end{tablenotes}
+\end{threeparttable}
 \end{table*}
@@
 As expected, for the same dimensionality a binary model achieves lower ranking accuracy. On average, the accuracy loss is around 8\%, varying between 2\% and 16\%.
-Nevertheless, the binary model's speed enables some very favourable tradeoffs
+Nevertheless, the model's speed enables some very favourable tradeoffs.
 

### 2017-06-13 paper/paper.tex
Offer instead of enable.

@@
 Large-scale recommender systems often face severe latency and storage constraints at prediction time.
-To relax those, we train recommendation models that use binary rather than real-valued user and item representations, and show that they enable a range of attractive speed-accuracy trade-offs.
+To relax those, we train recommendation models that use binary rather than real-valued user and item representations, and show that they offer a range of attractive speed-accuracy trade-offs.
 In particular, we demonstrate 13 times faster prediction at the expense of a modest 4\% loss in ranking accuracy on the Movielens 100K dataset.

### 2017-06-13 paper/paper.tex
Move related work to the end.
Small improvements.

@@
 To alleviate these constraints, we propose the use of binary item and user representations in learning-to-rank matrix factorization models.
-Following the approach of \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable ranking accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory.
+Following \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable ranking accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory.
 In MovieLens 100k experiments, our binary representations are within 4\% of the accuracy of the real-valued model, but are over 13 times faster to score and take 55\% of the memory.
+We make our model fitting and prediction code available on Github at \url{https://github.com/maciejkula/binge}.
 
-\section{Related work}
-\citet{shrivastava2014asymmetric} propose a solution for efficient ranking of large result sets through approximate maximum inner product search (MIPS).
-They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmetric transformations of the query and candidate vectors, and show their applicability to recommender systems.
+\section{The model}
+\label{sec:approach}
+XNOR-Networks \citep{rastegari2016xnor} combine 1-bit quantization of fully-connected and convolutional layers with representing dot products through scaled XNOR and bitcounting operations.
 
-Our approach is complementary to approximate inner product search.
-Approximate MIPS proceeds in two stages: (1) identification of candidate result vectors through localicty-sensitive hashing, followed by (2) explicit scoring and sorting of the returned candidates.
-Replacing real-valued dot products by XNOR operations in the second stage will yield incremental performance gains over a real-valued MIPS approach.
-
-\citet{Shen_2015_ICCV} expand on this approach by learning asymmetric binary hash functions.
-The hash functions are trained to minimize the L2 norm between the inner product matrix of the original input vectors and the inner product matrix of their binary representations.
-As a post-processing technique, it is a applicable to a wide range of problems.
-However, because the binary representations are obtained as a separate step, this approach may not be able to exploit the benefits in training a binary representation approach end-to-end.
-
-\section{Our approach}
-\label{sec:approach}
-\citet{rastegari2016xnor} introduce XNOR-Networks, an approach that combines 1-bit quantization of fully-connected and convolutional layers with representing dot products through scaled XNOR and bitcounting operations.
-
-We adapt their approach to the recommendation task by learning binary versions of user and item vectors within a learning-to-rank factorization model.
+We adapt them to the recommendation task by learning binary versions of user and item vectors within a learning-to-rank factorization model.
 This allows us to use binary instead of floating point operations for computing recommendation rankings, which has significant speed and memory use advantages.
@@
 In an equivalent binary model, $\frac{3}{32}n$ binary operations are required for each prediction (XOR, negation, and bitcounting), and the representations take $\frac{1}{8}n$ bytes of memory.
-Naive calculations suggest that, for the same latent dimensionality, a binary model would be over 20 times faster to score, and take less then 5\% of memory to store its representations.
+Na{\"\i}ve calculations suggest that, for the same latent dimensionality, a binary model would be over 20 times faster to score, and take less then 5\% of memory to store its representations.
 
@@
 Benchmark results are measured on an Intel Xeon E5-2686v4 CPU by running a 100 repetitions of scoring 100,000 items and averaging the results.
+Memory use is given by the total number of bytes required to store all the parameters of each model (including biases) and scaling parameters in the binary model.
 
@@
 The minibatch size and number of training epochs are treated as model hyperparameters.
-All models are trained using the Adam training rate schedule.
+All models are trained using the Adam training rate schedule \citep{kingma2014adam}.
 
@@
 SIMD (Single Instruction Multiple Data) instructions allow the CPU to operate on multiple pieces of data in parallel, achieving significant speedups over the scalar version.
-We use explicit intrinsics rather than compiler autovectorization to ensure that neither the real-valued nor the binary prediction code is unfairly disadvantaged by the quality of compiler autovectorization.
+We use explicit intrinsics rather than compiler autovectorization to ensure that neither the real-valued nor the binary prediction code is unfairly disadvantaged by the quality of automatically-generated SIMD operations.
 
@@
 
-The source code for the implementation is available on Github\footnote{\url{https://github.com/maciejkula/binge}}.
-
 \subsection{Results}
@@
 
-Results on scoring speed and memory use broadly confirm the naive calculations from section \ref{sec:approach}, converging to over 20 times faster scoring at around 3\% of memory as latent dimensionality increases.
+Results on scoring speed and memory use broadly confirm the na{\"\i}ve calculations from section \ref{sec:approach}, converging to over 20 times faster scoring at around 3\% of memory as latent dimensionality increases.
 As expected, for the same dimensionality a binary model achieves lower ranking accuracy. On average, the accuracy loss is around 8\%, varying between 2\% and 16\%.
-Nevertheless, the model's speed enables some very favourable tradeoffs.
 
+Nevertheless, the model's speed enables some favourable tradeoffs.
 When optimizing for MRR in both settings (at 1024 dimensions for the continuous and 2048 for the binary case), the XNOR representation achieves 96\% of the maximum MRR of the float representations at 13 times the query throughput and using only 6\% of the memory.
@@
 When optimizing for query throughput at 32 dimensions for the continuous case, the XNOR representation (at 512 dimensions) can improve the ranking performance by achieving 110\% of the continuous MRR while being slightly faster at 1.2 times the query throughput, and using 55\% of the memory.
+
+\section{Related work}
+\citet{shrivastava2014asymmetric} propose a solution for efficient ranking of large result sets through approximate maximum inner product search (MIPS).
+They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmetric transformations of the query and candidate vectors, and show their applicability to recommender systems.
+
+Our method is complementary to approximate inner product search.
+Approximate MIPS proceeds in two stages: (1) identification of candidate result vectors through localicty-sensitive hashing, followed by (2) explicit scoring and sorting of the returned candidates.
+Replacing real-valued dot products by XNOR operations in the second stage will yield incremental performance gains over a real-valued MIPS approach.
+
+\citet{Shen_2015_ICCV} expand on this approach by learning asymmetric binary hash functions.
+The hash functions are trained to minimize the L2 norm between the inner product matrix of the original input vectors and the inner product matrix of their binary representations.
+As a post-processing technique, it is a applicable to a wide range of problems.
+However, because the binary representations are obtained as a separate step, this approach may not be able to exploit the benefits in training a binary representation approach end-to-end.
 

### 2017-06-14 paper/paper.tex
Elaborate on what scale means. Add pinterest references.

--- 2017-06-13 paper/paper.tex
@@
 Large-scale recommender systems often face severe latency and storage constraints at prediction time.
-To relax those, we train recommendation models that use binary rather than real-valued user and item representations, and show that they offer a range of attractive speed-accuracy trade-offs.
+These are particularly acute when the number of items that could be recommended is large, and calculating predictions for the full set is computationally intensive.
+To relax these constraints, we train recommendation models that use binary rather than real-valued user and item representations, and show that they offer a range of attractive speed-accuracy trade-offs.
 In particular, we demonstrate 13 times faster prediction at the expense of a modest 4\% loss in ranking accuracy on the Movielens 100K dataset.
@@
 Industry ranking and recommendation systems need to scale to tens or hundreds of millions of items and users.
-Ranking latency (in on-line settings) and compute and storage costs (in pre-compute settings) place significant contraints on the design of the resulting system.
+As the number of items available to be recommended grows large, the challenge of etimating models from large amounts of interaction data is joined by the challenge of efficiently ranking the item catalogue to produce top-ranked recommendations.
+With large catalogues, ranking latency (in on-line settings) and compute and storage costs (in pre-compute settings) place significant contraints on the design of the entire recommender system.
 
@@
 Respecting that budget often means trading off model accuracy for speed, either via smaller and less expressive models (with lower-dimensionality representations) or through aggressive use of heuristics to exclude large classes of candidate items from scoring.
-For example, approximate nearest neighbour search techniques are used at YouTube \citep{covington2016deep} to reduce the number of candidate videos that need to be scored for each user.
+For example, Pinterest uses a combination of heuristics and candidate-generation models to select approximately 1000 pins, from billions of available pins, as input to its related pins ranking system \citep{liu2017related}.
+Similarly, approximate nearest neighbour search techniques are used at YouTube \citep{covington2016deep} to reduce the number of candidate videos that need to be scored for each user.
 

### 2017-06-14 paper/paper.tex
Exploratory changes.

--- 2017-06-13 paper/paper.tex
@@
 
-\section{Related work}
-\citet{shrivastava2014asymmetric} propose a solution for efficient ranking of large result sets through approximate maximum inner product search (MIPS).
+\section{Complementary and related approaches}
+Our work complements rather than replaces existing approaches to efficient ranking through approximate candidate selection.
+
+In particular, binary representations can be used in conjunction with approximate maximum inner product search (MIPS).
+\cite{shrivastava2014asymmetric} show that well-known locality-sensitive hashing techniques \citep{indyk1998approximate} can be extended to approximate MIPS through asymmetric transformations of the query and candidate vectors, leading to a two-step process whereby candidate vectors are identified through asymmetric hashing
+
+%% \section{Related work}
+%% Our model is complementary to methods that focus on restricting the set of candidates that need to be explicitly scored by the ranking system.
+
+%% Our model can be used in conjuction with methods that bound approximate nearest neighbours and maximum inner product approaches
+%% Our model is complementary to the chief \citet{shrivastava2014asymmetric} propose a solution for efficient ranking of large result sets through approximate maximum inner product search (MIPS).
 They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmetric transformations of the query and candidate vectors, and show their applicability to recommender systems.
 
-Our method is complementary to approximate inner product search.
+%% Our method is complementary to approximate inner product search.
 Approximate MIPS proceeds in two stages: (1) identification of candidate result vectors through localicty-sensitive hashing, followed by (2) explicit scoring and sorting of the returned candidates.
@@
 
-\citet{Shen_2015_ICCV} expand on this approach by learning asymmetric binary hash functions.
+%% \citet{Shen_2015_ICCV} expand on this approach by learning asymmetric binary hash functions.
 The hash functions are trained to minimize the L2 norm between the inner product matrix of the original input vectors and the inner product matrix of their binary representations.

### 2017-06-14 paper/paper.tex
Reformulate awkward sentence.

--- 2017-06-14 paper/paper.tex
@@
 Industry ranking and recommendation systems need to scale to tens or hundreds of millions of items and users.
-As the number of items available to be recommended grows large, the challenge of etimating models from large amounts of interaction data is joined by the challenge of efficiently ranking the item catalogue to produce top-ranked recommendations.
+As the number of items available to be recommended grows large, efficiently ranking the item catalogue to produce top-ranked recommendations becomes increasingly challenging.
 With large catalogues, ranking latency (in on-line settings) and compute and storage costs (in pre-compute settings) place significant contraints on the design of the entire recommender system.

### 2017-06-14 paper/paper.tex
More work on related section.

@@
 In MovieLens 100k experiments, our binary representations are within 4\% of the accuracy of the real-valued model, but are over 13 times faster to score and take 55\% of the memory.
-
-\section{Related work}
-\citet{shrivastava2014asymmetric} propose a solution for efficient ranking of large result sets through approximate maximum inner product search (MIPS).
-They extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to the MIPS setting through asymmetric transformations of the query and candidate vectors, and show their applicability to recommender systems.
-
-Our approach is complementary to approximate inner product search.
-Approximate MIPS proceeds in two stages: (1) identification of candidate result vectors through localicty-sensitive hashing, followed by (2) explicit scoring and sorting of the returned candidates.
-Replacing real-valued dot products by XNOR operations in the second stage will yield incremental performance gains over a real-valued MIPS approach.
-
-\citet{Shen_2015_ICCV} expand on this approach by learning asymmetric binary hash functions.
-The hash functions are trained to minimize the L2 norm between the inner product matrix of the original input vectors and the inner product matrix of their binary representations.
-As a post-processing technique, it is a applicable to a wide range of problems.
-However, because the binary representations are obtained as a separate step, this approach may not be able to exploit the benefits in training a binary representation approach end-to-end.
 
@@
 
+\section{Related work}
+Much of existing work on efficient recommendation retrieval focuses on reducing the number of dot products that need to be performed in order to retrieve top recommendations.
+\citet{koenigstein2012efficient} introduces an exact branch-and-bound based on metric trees as well as an approximate algorithm that clusters users and serves recommendations computed for the cluster centers.
+\citet{shrivastava2014asymmetric} extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to maximum inner product search (MIPS) through asymmetric transformations of the query and candidate vectors.
+
+Our work complements these approaches.
+Binary representations are direct replacement for existing real-valued recommendations, and will serve to speed up the final predictions on the set of candidates returned from tree nodes or hash buckets.
+
+The method that is most closely related to the model we present in this paper is \citet{Shen_2015_ICCV}, who expand on MIPS by learning asymmetric binary hash functions.
+The hash functions are trained to minimize the L2 norm between the inner product matrix of the original input vectors and the inner product matrix of their binary representations.
+As a post-processing technique, it is a applicable to a wide range of problems.
+However, because the binary representations are obtained as a separate step, this approach may not be able to exploit the benefits in training a binary representation approach end-to-end.
+
 \section{Conclusion}

### 2017-06-15 paper/paper.tex
Replace indicator function.
Explicitly mention usefulness in precompute settings.

@@
 This leads to often complex architectures that mix fairly simple on-line models, computationally expensive offline models, and intermediate complexity near-line algorithms where model updating is important \citep{amatriain2013big}.
+
 Additionaly, while a pre-compute solution removes model prediction from the critical path, it still requires substantial investment in computational and storage resources to calculate and cache the predictions of the model.
+Often the amount of comutation required makes timely updating of the cached results impossible.
+At Pinterest, pre-calculation of related pins results was so demanding that ranking of different segments of the catalogue had to be staggered in time, leading to stale results and reduced ease of development \citep[Section 5.5]{liu2017related}.
+\citet{koenigstein2012efficient} estimate that computing recommendations for the Yahoo!
+Music dataset \citep{dror2011yahoo} with a 50-dimensional latent factor model woud take over 135 hours.
 
-To alleviate these constraints, we propose the use of binary item and user representations in learning-to-rank matrix factorization models.
+To alleviate these constraints in both online and offline settings, we propose the use of binary item and user representations in learning-to-rank matrix factorization models.
 Following the approach of \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable ranking accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory.
@@
 
-The model is trained using backpropagation, and so we a smooth approximation to the derivative of the sign function to facilitate training. Following \citet{courbariaux2016binarized}, we define
+The model is trained using backpropagation, and so we use a smooth approximation to the derivative of the sign function to facilitate training. Following \citet{courbariaux2016binarized}, we define
 \begin{equation}
-\frac{d\sign}{dw} = w1_{\abs{w} \leq 1}.
+  \frac{d\sign}{dw}=
+  \begin{cases}
+    w & \text{if}\ \abs{w} \leq 1 \\
+    0 & \text{otherwise}.
+  \end{cases}
 \end{equation}

### 2017-06-15 paper/paper.tex
Fix memory use in introduction.

@@
 Following the approach of \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable ranking accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory.
-In MovieLens 100k experiments, our binary representations are within 4\% of the accuracy of the real-valued model, but are over 13 times faster to score and take 55\% of the memory.
+In MovieLens 100k experiments, our binary representations are within 4\% of the accuracy of the real-valued model, but are over 13 times faster to score and take only 6\% of the memory.
 

### 2017-06-16 paper/paper.tex
Fix footnote.
Stress that performance improvements are linear in the size of the catalogue.

@@
 The dataset contains 100 thousand ratings from 943 users on 1682 movies.
-Despite its small size, the results on this dataset should be indicative of results that can be obtained on larger datasets, and its small size enables us to easily test hyperparameter configurations and estimate models with high latent dimensions.
+Because the computational speed improvements offered by the binary representations are linear in the size of the catalogue, the results on this dataset should be indicative of results that can be obtained on larger datasets.
+At the same time, the dataset's small size enables us to easily test hyperparameter configurations and estimate models with high latent dimensions.
 
@@
 
-The source code for the implementation is available on Github\footnote{\url{https://github.com/maciejkula/binge}}.
+The source code for the implementation is available on Github at \url{https://github.com/maciejkula/binge}.
 

### 2017-06-16 binge/models.py
Gradient experiments.

@@ -45,14 +45,14 @@ def binarize_array(array):
     return array
 
 
-def binary_dot(x, y):
+# def binary_dot(x, y):
 
-    x_scale = x.abs().mean(1).detach()
-    y_scale = y.abs().mean(1).detach()
+#     x_scale = x.abs().mean(1)
+#     y_scale = y.abs().mean(1)
 
-    xnor = sign(x) * sign(y)
+#     xnor = sign(x) * sign(y)
 
-    return xnor.sum(1) * x_scale * y_scale
+#     return xnor.sum(1) * x_scale * y_scale
 
 
 class BinaryDot(Function):
@@ -77,23 +77,35 @@ class BinaryDot(Function):
 
         embedding_dim = x.size()[1]
 
-        x_scale = x.abs().mean(1)
-        y_scale = y.abs().mean(1)
+        grad_output = grad_output.expand_as(x)
 
-        sign_x = x.sign() * x_scale.expand_as(x)
-        sign_y = y.sign() * y_scale.expand_as(y)
+        x_scale = x.abs().mean(1).expand_as(x)
+        y_scale = y.abs().mean(1).expand_as(y)
 
-        return (sign_y * grad_output.expand_as(y)
-                * (1.0 / embedding_dim + (x.abs() < 1.0).float().abs()
-                   * x_scale.expand_as(y)),
-                sign_x * grad_output.expand_as(x)
-                * (1.0 / embedding_dim + (y.abs() < 1.0).float().abs()
-                   * y_scale.expand_as(x)))
+        sign_x = x.sign()
+        sign_y = y.sign()
 
+        dx_dsign = (x.abs() < 1.0).float() * x
+        dy_dsign = (y.abs() < 1.0).float() * y
 
-# def binary_dot(x, y):
+        # return (grad_output * sign_y *
+        #         (1.0 / embedding_dim + dx_dsign * x_scale),
+        #         grad_output * sign_x *
+        #         (1.0 / embedding_dim + dy_dsign * y_scale))
 
-#     return BinaryDot()(x, y)
+        return (grad_output * sign_y * y_scale *
+                (1.0 + dx_dsign * x_scale),
+                grad_output * sign_x * x_scale *
+                (1.0 + dy_dsign * y_scale))
+
+        return (grad_output * sign_y * y_scale *
+                (1.0 / embedding_dim + dx_dsign * x_scale),
+                grad_output * sign_x * x_scale *
[…]

### 2017-06-16 paper/paper.tex

@@
 \end{itemize}
-For both loss functions, for any known positive user-item interaction pair $(u, i)$, we sample an implicit negative item $j \in \symUserInteractionSet^-$.
+For both loss functions, for any known positive user-item interaction pair $(u, i)$, we uniformly sample an implicit negative item $j$ \in \symUserInteractionSet^-$.
 For BPR, the loss for any such triplet is given by
@@
 Results on scoring speed and memory use broadly confirm the naive calculations from section \ref{sec:approach}, converging to over 20 times faster scoring at around 3\% of memory as latent dimensionality increases.
+Memory savings plateau as latent dimensionality increases and storing the binary scaling factors becomes less important.
+
 As expected, for the same dimensionality a binary model achieves lower ranking accuracy. On average, the accuracy loss is around 8\%, varying between 2\% and 16\%.

### 2017-06-17 paper/paper.tex
Mention simplicity of implementation.

@@
 To alleviate these constraints in both online and offline settings, we propose the use of binary item and user representations in learning-to-rank matrix factorization models.
-Following the approach of \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable ranking accuracy to standard learning-to-rank approaches while being several times faster to score and requiring a fraction of the memory.
+Following the approach of \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable ranking accuracy to standard learning-to-rank approaches while being several times faster to score, requiring a fraction of the memory, and being only marginally more complex to estimate.
 In MovieLens 100k experiments, our binary representations are within 4\% of the accuracy of the real-valued model, but are over 13 times faster to score and take only 6\% of the memory.
@@
 \end{itemize}
-For both loss functions, for any known positive user-item interaction pair $(u, i)$, we uniformly sample an implicit negative item $j$ \in \symUserInteractionSet^-$.
+For both loss functions, for any known positive user-item interaction pair $(u, i)$, we uniformly sample an implicit negative item $j \in \symUserInteractionSet^-$.
 For BPR, the loss for any such triplet is given by
@@
 The model is implemented in PyTorch and trained using the nVidia K40 GPUs.
-During training, the embedding parameters are stored as floats.
+The implementation is only marginally more complex than a standard learning-to-rank model, and is accomplished by simply replacing the user-item vector dot product operation with its binary counterpart.
+The change results in a small increase in training time.
+
+During training, the embedding parameters are stored and updated as single precision floating point values.
 Similarly, the XNOR dot product is carried out using floats in $\{1, -1\}$.
 The minibatch size and number of training epochs are treated as model hyperparameters.
-All models are trained using the Adam training rate schedule.
+All models are trained using the Adam training rate schedule \citep{kingma2014adam}.
 
-The prediction code runs on the CPU and is implemented using Intel X86 AVX2 SIMD intrinsics.
+The prediction code runs on the CPU and is implemented in C using Intel X86 AVX2 SIMD intrinsics.
 SIMD (Single Instruction Multiple Data) instructions allow the CPU to operate on multiple pieces of data in parallel, achieving significant speedups over the scalar version.
@@
 The real-valued prediction code is implemented using 8-float wide fused multiply-add instructions (\texttt{\_mm256\_fmadd\_ps}).
-
 In the binary version, the real-valued embedding parameters used in training are discarded, and the derived 1-bit weights are packed into 32-bit integer buffers.
@@
 
-The source code for the implementation is available on Github at \url{https://github.com/maciejkula/binge}.
+The source code for both model training and prediction is available on Github at \url{https://github.com/maciejkula/binge}.
 
@@
 
-\section{Conclusion}
+\section{Conclusion and future work}
 We show that binary user and item representations allow recommender system designers to exploit an attractive new set of speed-accuracy trade-offs.

### 2017-06-17 binge/models.py
Change gradients.

@@ -45,16 +45,6 @@ def binarize_array(array):
     return array
 
 
-# def binary_dot(x, y):
-
-#     x_scale = x.abs().mean(1)
-#     y_scale = y.abs().mean(1)
-
-#     xnor = sign(x) * sign(y)
-
-#     return xnor.sum(1) * x_scale * y_scale
-
-
 class BinaryDot(Function):
 
     def forward(self, x, y):
@@ -85,47 +75,19 @@ class BinaryDot(Function):
         sign_x = x.sign()
         sign_y = y.sign()
 
-        dx_dsign = (x.abs() < 1.0).float() * x
-        dy_dsign = (y.abs() < 1.0).float() * y
+        dx_dsign = (x.abs() <= 1.0).float()
+        dy_dsign = (y.abs() <= 1.0).float()
 
-        # return (grad_output * sign_y *
-        #         (1.0 / embedding_dim + dx_dsign * x_scale),
-        #         grad_output * sign_x *
-        #         (1.0 / embedding_dim + dy_dsign * y_scale))
+        grads = (grad_output * sign_y * y_scale *
+                 (1.0 / embedding_dim + dx_dsign * x_scale),
+                 grad_output * sign_x * x_scale *
+                 (1.0 / embedding_dim + dy_dsign * y_scale))
 
-        return (grad_output * sign_y * y_scale *
-                (1.0 + dx_dsign * x_scale),
-                grad_output * sign_x * x_scale *
-                (1.0 + dy_dsign * y_scale))
-
-        return (grad_output * sign_y * y_scale *
-                (1.0 / embedding_dim + dx_dsign * x_scale),
-                grad_output * sign_x * x_scale *
-                (1.0 / embedding_dim + dy_dsign * y_scale))
+        return grads
 
 def binary_dot(x, y):
 
     return BinaryDot()(x, y)
-
-
-class Sign(Function):
-
-    def forward(self, x):
-
-        self.save_for_backward(x)
-
-        return x.sign()
-
-    def backward(self, grad_output):
-
-        x, = self.saved_tensors
-
-        return grad_output * (x.abs() < 1.0).float().abs()
-
-
-def sign(x):
-
-    return Sign()(x)
 
 
 class BilinearNet(nn.Module):

### 2017-06-17 paper/paper.tex
Add some future work.

@@
 
+There are a number of ways in which our work could be extended.
+An increasing number of recommender applications employ more sophisticated models using recurrent units for sequence representations or convolutional units for incorporating image data.
+Our approach can easily be applied in such models by simply replacing the topmost dot product or softmax layers with their binary equivalents.
+
 \bibliography{bibliography}

### 2017-06-19 binge/predict.c
Add lowdim simopts.

@@ -115,6 +115,72 @@ void predict_float_256(float* user_vector,
 }
 
 
+int predict_xnor_256_lowdim(int32_t* user_vector,
+                             int32_t* item_vectors,
+                             float user_bias,
+                             float* item_biases,
+                             float user_norm,
+                             float* item_norms,
+                             float* out,
+                             intptr_t num_items,
+                             intptr_t latent_dim,
+                             int cpuid) {
+    
+    int i, last_idx, item_idx;
+    int32_t* item_vector;
+    int total_elements = num_items * latent_dim;
+
+    __m256i x, y, xnor;
+    float scalar_prediction;
+    unsigned int on_bits;
+    int32_t bits[8] = {0, 0, 0, 0, 0, 0, 0, 0};
+
+    float max_on_bits = latent_dim * 32;
+
+    __m256i allbits = _mm256_cmpeq_epi32(
+        _mm256_setzero_si256(),
+        _mm256_setzero_si256());
+    
+    // Repeat the user vector to fit into a 256-bit AVX2 register
+    int32_t user_vector_repeated[8] = {0, 0, 0, 0, 0, 0, 0, 0};
+
+    for (int k; k < 8; k++) {
+        user_vector_repeated[k] = user_vector[k % latent_dim];
+    }
+
+    y = _mm256_loadu_si256(user_vector_repeated);
+
+    for (i=0; i + 8 < total_elements; i+=8) {
+
+        item_idx = i / latent_dim;
+
+        item_vector = item_vectors + i;
+
+        x = _mm256_loadu_si256(item_vector);
+
+        // XNOR
+        xnor = _mm256_xor_si256(_mm256_xor_si256(x, y), allbits);
+        _mm256_store_si256(bits, xnor);
+
+        // Bitcount
+        for (int k=0; k < 8 / latent_dim; k++) {
+
+            last_idx = item_idx + k;
+            
+            on_bits = popcnt_no_cpuid((const void*) (bits + k * latent_dim),
+                                      latent_dim * sizeof(float),
+                                      cpuid);
+            scalar_prediction = (on_bits - (max_on_bits - on_bits))
+                * item_norms[last_idx];
+                //* user_norm 
+            out[last_idx] = scalar_prediction + item_biases[last_idx];
+        }
+    }
+
+    return last_idx;
+}
+
+
 void predict_xnor_256(int32_t* user_vector,
                       int32_t* item_vectors,
                       float user_bias,
@@ -125,6 +191,22 @@ void predict_xnor_256(int32_t* user_vector,
[…]

### 2017-06-19 binge/data/movielens.py
Use movielens 1M

@@ -8,7 +8,14 @@ import scipy.sparse as sp
 from binge.data import _common
 
 
-def _read_raw_data(path):
+URL_100K = ('https://github.com/maciejkula/'
+            'lightfm_datasets/releases/'
+            'download/v0.1.0/movielens.zip')
+URL_10M = ('http://files.grouplens.org/datasets/movielens/ml-10m.zip')
+URL_1M = ('http://files.grouplens.org/datasets/movielens/ml-1m.zip')
+
+
+def _read_raw_data_100k(path):
     """
     Return the raw lines of the train and test files.
     """
@@ -17,19 +24,26 @@ def _read_raw_data(path):
         return datafile.read('ml-100k/u.data').decode().split('\n')
 
 
-def _parse(data):
+def _read_raw_data_10M(path):
+
+    with zipfile.ZipFile(path) as datafile:
+        # return datafile.read('ml-10M100K/ratings.dat').decode().split('\n')
+        return datafile.read('ml-1m/ratings.dat').decode().split('\n')
+
+
+def _parse(data, separator='\t'):
 
     for line in data:
 
         if not line:
             continue
 
-        uid, iid, rating, timestamp = [int(x) for x in line.split('\t')]
+        uid, iid, rating, timestamp = [x for x in line.split(separator)]
 
-        yield int(uid), int(iid), int(rating), int(timestamp)
+        yield int(uid), int(iid), float(rating), int(timestamp)
 
 
-def _load_data(data):
+def _load_data(data, separator='\t'):
 
     user_dict = {}
     item_dict = {}
@@ -38,7 +52,7 @@ def _load_data(data):
     iids = array.array('i')
     timestamps = array.array('i')
 
-    for uid, iid, rating, timestamp in _parse(data):
+    for uid, iid, rating, timestamp in _parse(data, separator):
         uids.append(user_dict.setdefault(uid,
                                          len(user_dict)))
         iids.append(item_dict.setdefault(iid,
@@ -91,7 +105,7 @@ def fetch_movielens(data_home=None, download_if_missing=True, random_seed=None):
                                 download_if_missing)
 
     # Load raw data
-    raw = _read_raw_data(zip_path)
+    raw = _read_raw_data_100k(zip_path)
 
     uids, iids, timestamps = _load_data(raw)
 
@@ -122,3 +136,70 @@ def fetch_movielens(data_home=None, download_if_missing=True, random_seed=None):
                           shape=(num_users, num_items))
 
[…]

### 2017-06-19 binge/predict.c

@@ -115,72 +115,6 @@ void predict_float_256(float* user_vector,
 }
 
 
-int predict_xnor_256_lowdim(int32_t* user_vector,
-                             int32_t* item_vectors,
-                             float user_bias,
-                             float* item_biases,
-                             float user_norm,
-                             float* item_norms,
-                             float* out,
-                             intptr_t num_items,
-                             intptr_t latent_dim,
-                             int cpuid) {
-    
-    int i, last_idx, item_idx;
-    int32_t* item_vector;
-    int total_elements = num_items * latent_dim;
-
-    __m256i x, y, xnor;
-    float scalar_prediction;
-    unsigned int on_bits;
-    int32_t bits[8] = {0, 0, 0, 0, 0, 0, 0, 0};
-
-    float max_on_bits = latent_dim * 32;
-
-    __m256i allbits = _mm256_cmpeq_epi32(
-        _mm256_setzero_si256(),
-        _mm256_setzero_si256());
-    
-    // Repeat the user vector to fit into a 256-bit AVX2 register
-    int32_t user_vector_repeated[8] = {0, 0, 0, 0, 0, 0, 0, 0};
-
-    for (int k; k < 8; k++) {
-        user_vector_repeated[k] = user_vector[k % latent_dim];
-    }
-
-    y = _mm256_loadu_si256(user_vector_repeated);
-
-    for (i=0; i + 8 < total_elements; i+=8) {
-
-        item_idx = i / latent_dim;
-
-        item_vector = item_vectors + i;
-
-        x = _mm256_loadu_si256(item_vector);
-
-        // XNOR
-        xnor = _mm256_xor_si256(_mm256_xor_si256(x, y), allbits);
-        _mm256_store_si256(bits, xnor);
-
-        // Bitcount
-        for (int k=0; k < 8 / latent_dim; k++) {
-
-            last_idx = item_idx + k;
-            
-            on_bits = popcnt_no_cpuid((const void*) (bits + k * latent_dim),
-                                      latent_dim * sizeof(float),
-                                      cpuid);
-            scalar_prediction = (on_bits - (max_on_bits - on_bits))
-                * item_norms[last_idx];
-                //* user_norm 
-            out[last_idx] = scalar_prediction + item_biases[last_idx];
-        }
-    }
-
-    return last_idx;
-}
-
-
 void predict_xnor_256(int32_t* user_vector,
                       int32_t* item_vectors,
                       float user_bias,
@@ -191,22 +125,6 @@ void predict_xnor_256(int32_t* user_vector,
[…]

### 2017-06-20 binge/models.py
Add popularity baseline. Bench with 500,000 items and take minimum.

@@ -483,3 +483,23 @@ class XNORScorer:
         get_size = lambda x: x.itemsize * x.size
 
         return sum(get_size(x) for x in self._parameters())
+
+
+class PopularityModel:
+
+    def __init__(self):
+
+        self._popularity = None
+
+    def fit(self, interactions):
+
+        self._popularity = interactions.getnnz(axis=0).astype(np.float32)
+
+        assert len(self._popularity) == interactions.shape[1]
+
+    def predict(self, user_ids, item_ids=None):
+
+        if item_ids is not None:
+            return self._popularity[item_ids]
+        else:
+            return self._popularity

### 2017-06-20 paper/paper.tex
Change names to use 1M dataset.

--- 2017-06-17 paper/paper.tex
@@
 To relax these constraints, we train recommendation models that use binary rather than real-valued user and item representations, and show that they offer a range of attractive speed-accuracy trade-offs.
-In particular, we demonstrate 13 times faster prediction at the expense of a modest 4\% loss in ranking accuracy on the Movielens 100K dataset.
+In particular, we demonstrate 13 times faster prediction at the expense of a modest 4\% loss in ranking accuracy on the Movielens 1M dataset.
 \end{abstract}
@@
 Following the approach of \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable ranking accuracy to standard learning-to-rank approaches while being several times faster to score, requiring a fraction of the memory, and being only marginally more complex to estimate.
-In MovieLens 100k experiments, our binary representations are within 4\% of the accuracy of the real-valued model, but are over 13 times faster to score and take only 6\% of the memory.
+In MovieLens 1M experiments, our binary representations are within 4\% of the accuracy of the real-valued model, but are over 13 times faster to score and take only 6\% of the memory.
 
@@
 \begin{threeparttable}
-\caption{Movielens 100K results}
+\caption{Movielens 1M results}
 \label{tb:results}
@@
 \end{table*}
-To assess the accuracy-speed trade-offs enabled by the XNOR approach, we conduct an experiment on the Movielens 100K dataset \citep{harper2016movielens}.
-The dataset contains 100 thousand ratings from 943 users on 1682 movies.
+To assess the accuracy-speed trade-offs enabled by the XNOR approach, we conduct an experiment on the Movielens 1M dataset \citep{harper2016movielens}.
+The dataset contains 1 million ratings from 6000 users on 4000 movies.
 Because the computational speed improvements offered by the binary representations are linear in the size of the catalogue, the results on this dataset should be indicative of results that can be obtained on larger datasets.
@@
 We randomly divide the dataset into a training, test, and validation set.
-In order to build a picture of the accuracy-speed trade-offs afforded by real-valued and binary latent models, we explicitly build models for between 32 and 4096 latent dimensions.
+In order to build a picture of the accuracy-speed trade-offs afforded by real-valued and binary latent models, we explicitly build models for between 32 and 2048 latent dimensions.
 For every latent dimensionality, we conduct a random search over the hyperparameter space, and pick optimal learning rates and minibatch sizes for each algorithm based on their performance on the test set.
@@
 \label{sec:results}
-Table \ref{tb:results} summarises the results of the Movielens 100K experiment.
+Table \ref{tb:results} summarises the results of the Movielens 1M experiment.
 Each row presents the best MRR (mean reciprocal rank) results for both real-valued and binary models of a given dimensionality.

### 2017-06-20 binge/data/movielens.py

@@ -11,7 +11,6 @@ from binge.data import _common
 URL_100K = ('https://github.com/maciejkula/'
             'lightfm_datasets/releases/'
             'download/v0.1.0/movielens.zip')
-URL_10M = ('http://files.grouplens.org/datasets/movielens/ml-10m.zip')
 URL_1M = ('http://files.grouplens.org/datasets/movielens/ml-1m.zip')
 
 
@@ -24,10 +23,9 @@ def _read_raw_data_100k(path):
         return datafile.read('ml-100k/u.data').decode().split('\n')
 
 
-def _read_raw_data_10M(path):
+def _read_raw_data_1M(path):
 
     with zipfile.ZipFile(path) as datafile:
-        # return datafile.read('ml-10M100K/ratings.dat').decode().split('\n')
         return datafile.read('ml-1m/ratings.dat').decode().split('\n')
 
 
@@ -64,7 +62,7 @@ def _load_data(data, separator='\t'):
             np.array(timestamps))
         
 
-def fetch_movielens(data_home=None, download_if_missing=True, random_seed=None):
+def fetch_movielens_100k(data_home=None, download_if_missing=True, random_seed=None):
     """
     Fetch the `Movielens 100k dataset <http://grouplens.org/datasets/movielens/100k/>`_.
 
@@ -138,7 +136,7 @@ def fetch_movielens(data_home=None, download_if_missing=True, random_seed=None):
     return train, test, validation
 
 
-def fetch_movielens_10M(data_home=None, download_if_missing=True, random_seed=None):
+def fetch_movielens_1M(data_home=None, download_if_missing=True, random_seed=None):
     """
     Parameters
     ----------
@@ -171,7 +169,7 @@ def fetch_movielens_10M(data_home=None, download_if_missing=True, random_seed=No
                                 download_if_missing)
 
     # Load raw data
-    raw = _read_raw_data_10M(zip_path)
+    raw = _read_raw_data_1M(zip_path)
 
     uids, iids, timestamps = _load_data(raw, '::')
 

### 2017-06-20 binge/rnn_data.py

@@ -1,148 +1,0 @@
-
-import zipfile
-
-import numpy as np
-
-import scipy.sparse as sp
-
-from lightfm.datasets import _common
-
-
-def _read_raw_data(path):
-    """
-    Return the raw lines of the train and test files.
-    """
-
-    with zipfile.ZipFile(path) as datafile:
-        return datafile.read('ml-100k/u.data').decode().split('\n')
-
-
-def _parse(data):
-
-    for line in data:
-
-        if not line:
-            continue
-
-        uid, iid, rating, timestamp = [int(x) for x in line.split('\t')]
-
-        yield uid, iid, rating, timestamp
-
-
-def _get_dimensions(data):
-
-    uids = set()
-    iids = set()
-    timestamps = set()
-
-    for uid, iid, _, timestamp in data:
-        uids.add(uid)
-        iids.add(iid)
-        timestamps.add(timestamp)
-
-    rows = max(uids) + 1
-    cols = max(iids) + 1
-    timestamps = max(timestamps) + 1
-
-    return rows, cols, timestamps
-
-
-def _build_interaction_matrices(rows, cols, data, min_rating):
-
-    train_mat = sp.lil_matrix((rows, cols), dtype=np.int32)
-    test_mat = sp.lil_matrix((rows, cols), dtype=np.int32)
-
-    uids_in_test = set()
-    uids_in_train = set()
-
-    for uid, iid, rating, timestamp in data:
-
-        if uid not in uids_in_test and uid not in uids_in_train:
-            if np.random.random() < 0.2:
-                uids_in_test.add(uid)
-            else:
-                uids_in_train.add(uid)
-
-            if uid in uids_in_test:
-                mat = test_mat
-            else:
-                mat = train_mat
-
-        if rating >= min_rating:
-            mat[uid, timestamp] = iid
-
-    return train_mat, test_mat
-
-
-def fetch_movielens(data_home=None, indicator_features=True, genre_features=False,
-                    min_rating=0.0, download_if_missing=True):
-    """
-    Fetch the `Movielens 100k dataset <http://grouplens.org/datasets/movielens/100k/>`_.
-
-    The dataset contains 100,000 interactions from 1000 users on 1700 movies,
-    and is exhaustively described in its
-    `README <http://files.grouplens.org/datasets/movielens/ml-100k-README.txt>`_.
-
-    Parameters
-    ----------
-
-    data_home: path, optional
-        Path to the directory in which the downloaded data should be placed.
[…]

### 2017-06-20 paper/paper.tex
Update with Movielens 1M results.

@@
 To relax these constraints, we train recommendation models that use binary rather than real-valued user and item representations, and show that they offer a range of attractive speed-accuracy trade-offs.
-In particular, we demonstrate 13 times faster prediction at the expense of a modest 4\% loss in ranking accuracy on the Movielens 1M dataset.
+In particular, we demonstrate 20 times faster prediction at the expense of a modest 7\% loss in ranking accuracy on the Movielens 1M dataset.
 \end{abstract}
@@
 \midrule
-       32 & 0.096 &      0.088 &     0.919 & 89,299 &     245,644 &      2.751 &            0.091 \\
-       64 & 0.098 &      0.096 &     0.973 & 47,565 &     209,486 &      4.404 &            0.062 \\
-      128 & 0.108 &      0.091 &     0.843 & 20,502 &     156,404 &      7.629 &            0.047 \\
-      256 & 0.109 &      0.097 &     0.884 & 10,631 &     166,500 &     15.661 &            0.039 \\
-      512 & 0.112 &      0.105 &     0.938 &   5391 &     110,418 &     20.479 &            0.035 \\
-     1024 & 0.114 &      0.102 &     0.897 &   2689 &      69,349 &     25.786 &            0.033 \\
-     2048 & 0.111 &      0.109 &     0.984 &   1352 &      34,826 &     25.745 &            0.032 \\
-     4096 & 0.111 &      0.107 &     0.961 &    666 &      15,907 &     23.856 &            0.032 \\
+       32 & 0.072 &      0.059 &     0.816 & 85,510 &     264,310 &      3.091 &            0.091 \\
+       64 & 0.075 &      0.065 &     0.862 & 44,964 &     221,000 &      4.915 &            0.062 \\
+      128 & 0.076 &      0.070 &     0.926 & 22,394 &     166,273 &      7.425 &            0.047 \\
+      256 & 0.078 &      0.072 &     0.916 & 12,040 &     189,353 &     15.727 &            0.039 \\
+      512 & 0.079 &      0.073 &     0.921 &   5933 &     118,895 &     20.039 &            0.035 \\
+     1024 & 0.081 &      0.076 &     0.933 &   2989 &      59,612 &     19.942 &            0.033 \\
+     2048 & 0.080 &      0.075 &     0.930 &   1514 &      32,092 &     21.185 &            0.032 \\
 \bottomrule
@@
 
-As expected, for the same dimensionality a binary model achieves lower ranking accuracy. On average, the accuracy loss is around 8\%, varying between 2\% and 16\%.
-Nevertheless, the model's speed enables some very favourable tradeoffs.
+As expected, for the same dimensionality a binary model achieves lower ranking accuracy. On average, the accuracy loss is around 10\%, varying between 7\% and 18\%.
+Nevertheless, the binary model's speed enables some very favourable tradeoffs
 
-When optimizing for MRR in both settings (at 1024 dimensions for the continuous and 2048 for the binary case), the XNOR representation achieves 96\% of the maximum MRR of the float representations at 13 times the query throughput and using only 6\% of the memory.
+When optimizing for MRR in both settings (at 1024 dimensions for the continuous and 1024 for the binary case), the XNOR representation achieves 93\% of the maximum MRR of the float representations at 20 times the query throughput and using only 3\% of the memory.
 
-When optimizing for query throughput at 32 dimensions for the continuous case, the XNOR representation (at 512 dimensions) can improve the ranking performance by achieving 110\% of the continuous MRR while being slightly faster at 1.2 times the query throughput, and using 55\% of the memory.
+When optimizing for query throughput at 32 dimensions for the continuous case, the XNOR representation (at 512 dimensions) can improve the ranking performance by achieving 101\% of the continuous MRR while being slightly faster at 1.4 times the query throughput, and using 55\% of the memory.
 

### 2017-06-21 paper/paper.tex
Revise and shorten paper.

@@
 \title{Binary Latent Representations for Efficient Ranking}
+\subtitle{Empirical Assessment}
 \author{Maciej Kula}
@@
 These are particularly acute when the number of items that could be recommended is large, and calculating predictions for the full set is computationally intensive.
-To relax these constraints, we train recommendation models that use binary rather than real-valued user and item representations, and show that they offer a range of attractive speed-accuracy trade-offs.
-In particular, we demonstrate 20 times faster prediction at the expense of a modest 7\% loss in ranking accuracy on the Movielens 1M dataset.
+In an attempt to relax these constraints, we train recommendation models that use binary rather than real-valued user and item representations, and show that while they are substantially faster to evaluate, the gains in speed come at a large cost in accuracy.
+In our Movielens 1M experiments, we show that reducing the latent dimensionality of traditional models offers a more attractive accuracy/speed trade-off than using binary representations.
 \end{abstract}
@@
 
-To alleviate these constraints in both online and offline settings, we propose the use of binary item and user representations in learning-to-rank matrix factorization models.
-Following the approach of \citet{rastegari2016xnor}, we show that we can estimate binary user and item representations that achieve comparable ranking accuracy to standard learning-to-rank approaches while being several times faster to score, requiring a fraction of the memory, and being only marginally more complex to estimate.
-In MovieLens 1M experiments, our binary representations are within 4\% of the accuracy of the real-valued model, but are over 13 times faster to score and take only 6\% of the memory.
+To alleviate these constraints in both online and offline settings, we evaluate the use of binary item and user representations in learning-to-rank matrix factorization models.
+Following the approach of \citet{rastegari2016xnor}, we estimate binary user and item representations that are several times faster to score and requiring a fraction of the memory (for the same latent dimensionality).
+Unfortunately, they are substantially less accurate than standard learning-to-rank approaches.
+The resulting speed-accuracy trade-offs favour simply reducing the latent dimensionality of traditional models when evaluation speed is desired.
 
-\section{Our approach}
+We speculate that binary representations are more useful in case of more complex models (such as deep convolutional models) where a single prediction requires many more floating operations than relatively compact biliearn recommender models.
+
+%%_______________________________________
+%% In MovieLens 1M experiments, our binary representations are within 4\% of the accuracy of the real-valued model, but are over 13 times faster to score and take only 6\% of the memory.
+%% ____________________________________
+
+\section{Binary latent representations}
 \label{sec:approach}
@@
 If accuracy is the priority, a binary model may offer equivalent ranking performance while being faster to score.
-This suspicion is borne out in our experiments.
 Secondly, many production systems resort to heavy use of heuristics when selecting candidates to be scored by their ranking systems (as scoring all candidates would be infeasible).
@@
 Note that the scaling factors $\alpha$ and $\beta$ need to be stored (in addition to the latent representations and biases) to compute predictions in the binary model.
+Their use also implies a constant cost of two floating-point multiplications per dot product when using binary representations.
 
@@
 In particular, models using recurrent \citep{hidasi2015session} or convolutional \citep{lynch2015images} item or user representations can easily be augmented to use binary dot products in the final ranking stages.
-Assuming item representations can be computed in advance, even very complicated models can enjoy the efficiency advantages of using binary representations.
 
@@
 \midrule
-       32 & 0.072 &      0.059 &     0.816 & 85,510 &     264,310 &      3.091 &            0.091 \\
-       64 & 0.075 &      0.065 &     0.862 & 44,964 &     221,000 &      4.915 &            0.062 \\
-      128 & 0.076 &      0.070 &     0.926 & 22,394 &     166,273 &      7.425 &            0.047 \\
-      256 & 0.078 &      0.072 &     0.916 & 12,040 &     189,353 &     15.727 &            0.039 \\
-      512 & 0.079 &      0.073 &     0.921 &   5933 &     118,895 &     20.039 &            0.035 \\
-     1024 & 0.081 &      0.076 &     0.933 &   2989 &      59,612 &     19.942 &            0.033 \\
-     2048 & 0.080 &      0.075 &     0.930 &   1514 &      32,092 &     21.185 &            0.032 \\
+       32 & 0.077 &      0.059 &     0.768 & 87,758 &     264,146 &      3.010 &            0.091 \\
+       64 & 0.075 &      0.066 &     0.878 & 45,992 &     220,833 &      4.802 &            0.062 \\
+      128 & 0.076 &      0.071 &     0.935 & 23,870 &     166,252 &      6.965 &            0.047 \\
+      256 & 0.078 &      0.071 &     0.908 & 12,284 &     190,131 &     15.477 &            0.039 \\
+      512 & 0.080 &      0.073 &     0.912 &   6074 &     122,637 &     20.190 &            0.035 \\
+     1024 & 0.080 &      0.075 &     0.936 &   3056 &      61,301 &     20.056 &            0.033 \\
 \bottomrule
@@
 We randomly divide the dataset into a training, test, and validation set.
-In order to build a picture of the accuracy-speed trade-offs afforded by real-valued and binary latent models, we explicitly build models for between 32 and 2048 latent dimensions.
-For every latent dimensionality, we conduct a random search over the hyperparameter space, and pick optimal learning rates and minibatch sizes for each algorithm based on their performance on the test set.
+In order to build a picture of the accuracy-speed trade-offs afforded by real-valued and binary latent models, we explicitly build models for between 32 and 1024 latent dimensions.
+For every latent dimensionality, we conduct a random search over the hyperparameter space, and pick optimal initial learning rate, loss function, L2 penalty, minibatch size and number of training epochs for each algorithm based on their performance on the test set.
 The final results are obtained from ranking interactions from the validation set.
@@
 
-Benchmark results are measured on an Intel Xeon E5-2686v4 CPU by running a 100 repetitions of scoring 100,000 items and averaging the results.
+Benchmark results are measured on an Intel Xeon E5-2686v4 CPU by running a 500 repetitions of scoring 100,000 items and averaging the results.
 
@@
 Similarly, the XNOR dot product is carried out using floats in $\{1, -1\}$.
-The minibatch size and number of training epochs are treated as model hyperparameters.
+The initial learning rate, loss function, L2 penalty, minibatch size and number of training epochs are treated as model hyperparameters.
 All models are trained using the Adam training rate schedule \citep{kingma2014adam}.
@@
 
-As expected, for the same dimensionality a binary model achieves lower ranking accuracy. On average, the accuracy loss is around 10\%, varying between 7\% and 18\%.
-Nevertheless, the binary model's speed enables some very favourable tradeoffs
+As expected, for the same dimensionality a binary model achieves lower ranking accuracy.
+On average, the accuracy loss when moving from a contious to a binary model of the same latent dimensionality is around 11\%, varying between 6\% and 23\%.
 
-When optimizing for MRR in both settings (at 1024 dimensions for the continuous and 1024 for the binary case), the XNOR representation achieves 93\% of the maximum MRR of the float representations at 20 times the query throughput and using only 3\% of the memory.
+Unfortunately, while continous models retain good accuracy as latent dimensionality decreases, binary models' representational power sharply deteriorates.
+Moving from the 1024 to 32 dimensions in the continuous model implies a 29 times increase in prediction speed at the expense of a modest 4\% decrease in accuracy.
+This compares favourably to switching to binary representations: moving to a 1024-binary dimensional representation implies a sharper accuracy drop at 6\% in exchange for a smaller 20 times increase in prediction speed.
+Moving to 32 dimensions yields a further speed gains at 86 times, but at the cost of a considerable loss of accuracy at 26\%.
 
-When optimizing for query throughput at 32 dimensions for the continuous case, the XNOR representation (at 512 dimensions) can improve the ranking performance by achieving 101\% of the continuous MRR while being slightly faster at 1.4 times the query throughput, and using 55\% of the memory.
+The results suggest that latent recommender models have good representational power even at low latent dimensionalities.
+This advantage is lost when using binary embeddings, which need to be high-dimensional to achieve comparable accuracy.
+At the same time, binary representations' speed advantage is only evident at high dimensions: at low dimensions, only floating-point operations can use enjoy the advantages of SIMD operations, and the fixed overhead of binary scaling factors $\alpha$ and $\beta$ constitutes a larger proportion of the total computational cost.
+We conjecture that the attractiveness of binary representations is tightly coupled with high-dimensional models where a single prediction requires many floating point operations, such as convolutional neural networks.
+Relatively compact latent factor models do not fall into this category.
 
-\section{Related work}
-Much of existing work on efficient recommendation retrieval focuses on reducing the number of dot products that need to be performed in order to retrieve top recommendations.
-\citet{koenigstein2012efficient} introduces an exact branch-and-bound based on metric trees as well as an approximate algorithm that clusters users and serves recommendations computed for the cluster centers.
-\citet{shrivastava2014asymmetric} extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to maximum inner product search (MIPS) through asymmetric transformations of the query and candidate vectors.
-
-Our work complements these approaches.
-Binary representations are direct replacement for existing real-valued recommendations, and will serve to speed up the final predictions on the set of candidates returned from tree nodes or hash buckets.
-
-The method that is most closely related to the model we present in this paper is \citet{Shen_2015_ICCV}, who expand on MIPS by learning asymmetric binary hash functions.
-The hash functions are trained to minimize the L2 norm between the inner product matrix of the original input vectors and the inner product matrix of their binary representations.
-As a post-processing technique, it is a applicable to a wide range of problems.
-However, because the binary representations are obtained as a separate step, this approach may not be able to exploit the benefits in training a binary representation approach end-to-end.
-
-\section{Conclusion and future work}
-We show that binary user and item representations allow recommender system designers to exploit an attractive new set of speed-accuracy trade-offs.
-They are straightforward to estimate, and can easily be used in any existing model estimated using stochastic gradient descent.
+\section{Conclusion}
+While prediction latency is a pressing concern for many recommender systems, we find that the already compact latent factor models do not stand to benefit from using binary latent representations.
+The sharp drop in representational power exhibited by binary embeddings makes other approaches (such as exact \citep{koenigstein2012efficient} and approximate \citet{shrivastava2014asymmetric} maximum inner product search) more promising avenues when optimizing large-scale ranking.
 

### 2017-06-22 paper/paper.tex
Spelling fixes.

@@
 
-We speculate that binary representations are more useful in case of more complex models (such as deep convolutional models) where a single prediction requires many more floating operations than relatively compact biliearn recommender models.
-
-%%_______________________________________
-%% In MovieLens 1M experiments, our binary representations are within 4\% of the accuracy of the real-valued model, but are over 13 times faster to score and take only 6\% of the memory.
-%% ____________________________________
+We speculate that binary representations are more useful in case of more complex models (such as deep convolutional models) where a single prediction requires many more floating operations than relatively compact bilinear recommender models.
 
@@
 Let $\symUserSet$ be the set of users, and $\symItemSet$ be the set of items.
-Each user interacts with a number of items $\symUserInteractionSet_i^+$; the set of all remaining items is denoted by $\symUserInteractionSet^-$.
+Each user interacts with a number of items $\symUserInteractionSet^+$; the set of all remaining items is denoted by $\symUserInteractionSet^-$.
 
@@
   \begin{cases}
-    w & \text{if}\ \abs{w} \leq 1 \\
+    1 & \text{if}\ \abs{w} \leq 1 \\
     0 & \text{otherwise}.
@@
 While prediction latency is a pressing concern for many recommender systems, we find that the already compact latent factor models do not stand to benefit from using binary latent representations.
-The sharp drop in representational power exhibited by binary embeddings makes other approaches (such as exact \citep{koenigstein2012efficient} and approximate \citet{shrivastava2014asymmetric} maximum inner product search) more promising avenues when optimizing large-scale ranking.
+The sharp drop in representational power exhibited by binary embeddings makes other approaches (such as exact \citep{koenigstein2012efficient} and approximate \citep{shrivastava2014asymmetric} maximum inner product search) more promising avenues when optimizing large-scale ranking.
 

### 2017-08-14 paper/paper.tex
Update table notes.

--- 2017-06-22 paper/paper.tex
@@
 \toprule
-Dimension &   MRR & Binary MRR & MRR ratio\tnote{1} &   PPMS\tnote{2} & Binary PPMS & PPMS ratio & Memory use ratio \\
+Dimension &   MRR & Binary MRR & MRR ratio\tnote{1} &   PPMS\tnote{2} & Binary PPMS & PPMS ratio\tnote{3} & Memory use ratio\tnote{4} \\
 \midrule
@@
 \item[1] Ratio of binary model MRR to real-valued model MRR
-\item[2] Predictions Per Millisecond
+\item[2] Predictions Per Millisecond: how many items can be scored per millisecond
+\item[3] Ratio of binary PPMS to real-valued PPMS
+\item[4] Ratio of memory required to store binary vs. real-valued parameters
 }

### 2017-08-14 paper/paper.tex
Add an alternatives section.

@@
 
+\section{Alternative approaches}
+Given the limited success of binary embedding, a more promising approach is offered by existing work that focuses on reducing the number of dot products which need to be performed in order to retrieve top recommendations.
+\citet{koenigstein2012efficient} introduces an exact branch-and-bound based on metric trees as well as an approximate algorithm that clusters users and serves recommendations computed for the cluster centers.
+\citet{shrivastava2014asymmetric} extend well-known \citep{indyk1998approximate} locality-sensitive hashing techniques to maximum inner product search (MIPS) through asymmetric transformations of the query and candidate vectors.
+
+The method that is most closely related to the model we present in this paper is \citet{Shen_2015_ICCV}, who expand on MIPS by learning asymmetric binary hash functions.
+The hash functions are trained to minimize the L2 norm between the inner product matrix of the original input vectors and the inner product matrix of their binary representations.
+
 \section{Conclusion}

### 2017-08-14 paper/paper.tex
Fix typos.

@@
 As the number of items available to be recommended grows large, efficiently ranking the item catalogue to produce top-ranked recommendations becomes increasingly challenging.
-With large catalogues, ranking latency (in on-line settings) and compute and storage costs (in pre-compute settings) place significant contraints on the design of the entire recommender system.
+With large catalogues, ranking latency (in on-line settings) and compute and storage costs (in pre-compute settings) place significant constraints on the design of the entire recommender system.
 
@@
 
-Additionaly, while a pre-compute solution removes model prediction from the critical path, it still requires substantial investment in computational and storage resources to calculate and cache the predictions of the model.
-Often the amount of comutation required makes timely updating of the cached results impossible.
+Additionally, while a pre-compute solution removes model prediction from the critical path, it still requires substantial investment in computational and storage resources to calculate and cache the predictions of the model.
+Often the amount of computation required makes timely updating of the cached results impossible.
 At Pinterest, pre-calculation of related pins results was so demanding that ranking of different segments of the catalogue had to be staggered in time, leading to stale results and reduced ease of development \citep[Section 5.5]{liu2017related}.
@@
 
-To quantify the tradeoffs involved, we construct and fit both a real-valued and a binary version of simple implicit learning-to-rank factorization model.
+To quantify the trade-offs involved, we construct and fit both a real-valued and a binary version of simple implicit learning-to-rank factorization model.
 We introduce the notation and the models below.
@@
 As expected, for the same dimensionality a binary model achieves lower ranking accuracy.
-On average, the accuracy loss when moving from a contious to a binary model of the same latent dimensionality is around 11\%, varying between 6\% and 23\%.
+On average, the accuracy loss when moving from a continuous to a binary model of the same latent dimensionality is around 11\%, varying between 6\% and 23\%.
 
-Unfortunately, while continous models retain good accuracy as latent dimensionality decreases, binary models' representational power sharply deteriorates.
+Unfortunately, while continuous models retain good accuracy as latent dimensionality decreases, binary models' representational power sharply deteriorates.
 Moving from the 1024 to 32 dimensions in the continuous model implies a 29 times increase in prediction speed at the expense of a modest 4\% decrease in accuracy.

### 2017-08-17 paper/paper.tex
Add keywords.

@@
 \acmConference[]{}{}{}
+
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10002951.10003260.10003261.10003269</concept_id>
+<concept_desc>Information systems~Collaborative filtering</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10002951.10003317.10003338.10003346</concept_id>
+<concept_desc>Information systems~Top-k retrieval in databases</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Information systems~Collaborative filtering}
+\ccsdesc[500]{Information systems~Top-k retrieval in databases}
+
+\keywords{Recommender Systems, Matrix Factorization, Binary Representations}
 
